### 2017-05-11 upi.tex
fixed pi2

\documentclass{entcs}
\usepackage{prentcsmacro}

\def\lastname{Carette, Chen, Choudhury, Rose, and Sabry}

%% Amr
%% words to remember :-)
%% sublime unfathomable
%% path categorical semantics
%% ---

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}

\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\texttt{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\texttt{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}

\newcommand{\Typ}{\mathbf{Type}}
\newcommand{\alt}{~\mid~}
\newcommand{\patht}[1]{\textsc{PATHS}(#1,#1)}
\newcommand{\fpatht}[1]{\textsc{FREEPATHS}(#1,\Box)}
\newcommand{\fpathp}[2]{\textsc{freepath}~#1~#2}
\newcommand{\pathind}[2]{\textsc{pathind}~#1~#2}
\newcommand{\invc}[1]{!\;#1}
\newcommand{\evalone}[2]{eval(#1,#2)}
\newcommand{\evalbone}[2]{evalB(#1,#2)}
\newcommand{\reflp}{\textsc{refl}}
\newcommand{\notp}{\textsc{not}}
\newcommand{\gluep}{\textsc{glue}}
\newcommand{\reflh}{\mathit{refl}_{\sim}}
\newcommand{\symh}[1]{\mathit{sym}_{\sim}~#1}
\newcommand{\transh}[2]{\mathit{trans}_{\sim}~#1~#2}
\newcommand{\reflq}{\mathit{refl}_{\simeq}}
\newcommand{\symq}[1]{\mathit{sym}_{\simeq}~#1}
\newcommand{\transq}[2]{\mathit{trans}_{\simeq}~#1~#2}
\newcommand{\isequiv}[1]{\mathit{isequiv}(#1)}
\newcommand{\idc}{\mathit{id}_{\boolt}}
\newcommand{\swapc}{\mathit{swap}_{\boolt}}
\newcommand{\assocc}{\mathit{assoc}}
\newcommand{\invl}{\mathit{invl}}
\newcommand{\invr}{\mathit{invr}}
\newcommand{\invinv}{\mathit{inv}^2}
\newcommand{\idlc}{\mathit{idl}}
\newcommand{\idrc}{\mathit{idr}}
\newcommand{\swapswap}{\swapc^2}
\newcommand{\compsim}{\compc_{\isotwo}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\isotwo}{\Leftrightarrow}
\newcommand{\piso}{\multimapdotbothB~~}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\bt}{\mathbb{2}}
\newcommand{\fc}{\mathit{false}}
\newcommand{\tc}{\mathit{true}}
\newcommand{\boolt}{\mathbb{B}}
\newcommand{\univ}{\mathcal{U}}
\newcommand{\uzero}{\mathcal{U}_0}
\newcommand{\uone}{\mathcal{U}_1}
\newcommand{\Rule}[2]{
\makebox{
$\displaystyle
\frac{\begin{array}{l}#1\\\end{array}}
{\begin{array}{l}#2\\\end{array}}$}}
\newcommand{\proves}{\vdash}
\newcommand{\jdgg}[3]{#1 \proves #2 : #3}
\newcommand{\jdg}[2]{\proves #1 : #2}
\newcommand{\jdge}[3]{\proves #1 = #2 : #3}
%% codes
%% denotations

\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Amr says: {#1}}\end{minipage}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frontmatter}
\title{From Reversible Programs to \\ Univalent Universes and Back}
\author{Jacques Carette}
\address{McMaster University}
\author{Chao-Hong Chen}
\address{Indiana University}
\author{Vikraman Choudhury}
\author{Robert Rose}
\author{Amr Sabry}
\address{Indiana University}

\begin{abstract}
  We establish a close connection between a reversible programming language
  based on type isomorphisms and a formally presented univalent universe. The
  correspondence relates combinators witnessing type isomorphisms in the
  programming language to paths in the univalent universe, and combinator
  optimizations in the programming language to 2-paths in the univalent
  universe. The result suggests a simple computational interpretation of paths
  and of univalence in terms of familiar programming constructs whenever the
  universe in question is computable.
\end{abstract}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

There is a cottage industry of reversible programming languages, reversible
logic, programming applications of type isomorphisms, etc. This work seems it
should be connected to HoTT and univalence but there aren't any precise
connections or theorems.

We propose a featherweight very specialized version of HoTT: a smaller
collection of types and only reversible functions. No dependent types!  More
precisely, our conjecture:

\begin{itemize}
\item Let $\mathcal{U}$ be the univalent subuniverse generated by $0$, $1$, $+$,
  and $*$ suitably truncated.
\item Let $\Pi$ be the previously  described reversible programming with 1-paths and 2-paths
\item We conjecture that $\Pi$ includes codes for \emph{all} paths in $\mathcal{U}$
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Simple Reversible Programming Language}

The practice of programming languages is replete with \emph{ad hoc} instances of
reversible computations such as database transactions, mechanisms for data
provenance, checkpoints, stack and exception traces, logs, backups, rollback
recoveries, version control systems, reverse engineering, software transactional
memories, continuations, backtracking search, and multiple-level ``undo''
features in commercial applications. In the early nineties,
Baker~\cite{Baker:1992:LLL,Baker:1992:NFT} argued for a systematic, first-class,
treatment of reversibility but intensive research in full-fledged reversible
models of computations and reversible programming languages was only sparked by
the discovery of deep connections between physics and
computation~\cite{Landauer:1961,PhysRevA.32.3266,Toffoli:1980,bennett1985fundamental,Frank:1999:REC:930275}
and by the potential for efficient quantum
computation~\cite{springerlink:10.1007/BF02650179}.

The early developments of reversible programming languages started with a
conventional programming language, e.g., an extended $\lambda$-calculus, and either
(i) extended the language with a history
mechanism~\cite{vanTonder:2004,Kluge:1999:SEMCD,lorenz,danos2004reversible} or
(ii) imposed constraints on the control flow constructs to make them
reversible~\cite{Yokoyama:2007:RPL:1244381.1244404}. More modern approaches
recognize that reversible programming languages require a fresh approach and
should be designed from first principles without the detour via conventional
irreversible languages~\cite{Yokoyama:2008:PRP,Mu:2004:ILRC,abramsky2005structural,DiPierro:2006:RCL:1166042.1166047}.

%%%%%
\subsection{The $\Pi$ Family of Languages}

In previous work, Carette, Bowman, James, and
Sabry~\cite{rc2011,James:2012:IE:2103656.2103667,Carette2016} introduced
the~$\Pi$ family of reversible languages based on type isomorphisms and
commutative semiring identities. The fragment without recursive types is
universal for reversible boolean circuits~\cite{James:2012:IE:2103656.2103667}
and the extension with recursive types and trace
operators~\cite{Hasegawa:1997:RCS:645893.671607} is a Turing-complete reversible
language~\cite{James:2012:IE:2103656.2103667,rc2011}. In this paper we will
identify a ``featherweight'' representative to use for the formal
development. But first, we illustrate the flavor of the family of languages with
some examples written in a Haskell syntax. First consider the following program:

{\footnotesize
\begin{verbatim}
addSub1 :: (Int, Int) :<=> (Int, Int)
addSub1 = CommuteTimes
      :.: (UnfoldN :*: Id)
      :.: Distribute
      :.: (Id :+: CommuteTimes)
      :.: Factor
      :.: (FoldN :*: Id)
\end{verbatim}}
%
\noindent which takes a pair of natural numbers and processes them as follows depending on
whether the second number is 0 or not. If the second number is 0, the evaluation
proceeds as follows:
$(n,0) \mapsto (0,n) \mapsto (\mathit{inj}_1(),n) \mapsto \mathit{inj}_1((),n)
\mapsto \mathit{inj}_1((),n) \mapsto (\mathit{inj}_1(),n) \mapsto (0,n)$. If the
second number is $\mathit{suc}~m$, the evaluation proceeds as follows:
$(n, \mathit{suc}~m) \mapsto (\mathit{suc}~m,n) \mapsto (\mathit{inj}_2m,n)
\mapsto \mathit{inj}_2(m,n) \mapsto \mathit{inj}_2(n,m) \mapsto
(\mathit{inj}_2n,m) \mapsto (\mathit{suc}~n,m)$. In summary, repeated
applications of this program would increment the first number and decrement the
second number maintaining the sum at all times. When the second number reaches
0, the numbers are swapped.

As another example, consider the following expressions culminating with the
implementation of the Toffoli gate, which is universal for combinational
circuits. We have two implementations of simple boolean negation: a direct one
and a more convoluted one. These two implementations are ``equivalent'' in the
sense to be explained below. The expression \verb|cond| is the main reversible
control flow mechanism: it allows for the conditional execution of one of two
branches but retains the choice of which branch was executed. This mechanism can
be used twice to implement the Toffoli gate, which performs a conditional
negation on one bit if two other bits are both true:

{\footnotesize
\begin{verbatim}
inot :: Bool :<=> Bool
inot = CommutePlus

inot2 :: Bool :<=> Bool
inot2 = UnitTimesI
    :.: CommuteTimes
    :.: (CommutePlus :*: Id)
    :.: CommuteTimes
    :.: UnitTimesE

cond :: (a :<=> b) -> (a :<=> b) -> ((Bool, a) :<=> (Bool, b))
cond f g = Distribute
       :.: ((Id :*: f) :+: (Id :*: g))
       :.: Factor

controlled :: (a :<=> a) -> ((Bool, a) :<=> (Bool, a))
controlled f = cond f Id

cnot :: (Bool, Bool) :<=> (Bool, Bool)
cnot = controlled inot

toffoli :: ((Bool,Bool),Bool) :<=> ((Bool,Bool),Bool)
toffoli = AssocTimesR :.: controlled cnot :.: AssocTimesL
\end{verbatim}}

Although writing circuits using the raw syntax for combinators above is tedious,
the examples illustrate the programming language nature of this family of $\Pi$
languages. In other work, one can find a compiler from a conventional functional
language to circuits~\cite{James:2012:IE:2103656.2103667}, a systematic
technique to translate abstract machines to $\Pi$~\cite{rc2012}, and a
Haskell-like surface language~\cite{theseus} which can ease writing
programs. All that reinforces the claim that this is a practical approach to
reversible programming.

The raw syntax for combinators has, however, a significant benefit: it allows
algebraic manipulation of the programs using a second level of combinators. As
an example it is possible to show the the two implementations of boolean
negation are equivalent as follows:

{\footnotesize
\begin{verbatim}
negOpt :: inot2 :<===> inot
negOpt = (Id :>: AssocLeftEquiv)
    :..: (Id :>: (SlideCommuteTimes :>: Id))
    :..: (Id :>: AssocRightEquiv)
    :..: (Id :>: (Id :>: AssocLeftEquiv))
    :..: (Id :>: (Id :>: (InvInvEquivId :>: Id)))
    :..: (Id :>: (Id :>: IdElimL))
    :..: AssocLeftEquiv
    :..: (SlideUnitI :>: Id)
    :..: AssocRightEquiv
    :..: (Id :>: InvInvEquivId)
    :..: IdElimR
\end{verbatim}}
%
\noindent This program, which can be thought of as an optimization or a proof, relates the
two implementations of boolean negation via a series of elementary
optimizations of which the following are the most interesting:

{\footnotesize
\begin{verbatim}
AssocLeftEquiv    ::           (c1 :.: (c2 :.: c3)) :<===> ((c1 :.: c2) :.: c3)
SlideCommuteTimes :: (CommuteTimes :.: (c1 :*: c2)) :<===> ((c2 :*: c1) :.: CommuteTimes)
IdElimL           ::                     (Id :.: c) :<===> c
SlideUnitI        ::   (UnitTimesI :.: (c1 :*: c2)) :<===> (c2 :.: UnitTimesI)
\end{verbatim}}

%%%%%
\subsection{$\Pi_{\mathbb{2}}$}

In previous work, on $\Pi$ we noted a possible connection with HoTT:
\begin{quote}
  It is, therefore, at least plausible that a variant of HoTT based exclusively
  on reversible functions that directly correspond to equivalences would have
  better computational properties. Our current result is a step, albeit
  preliminary, in that direction as it only applies to finite
  types~\cite{Carette2016}.
\end{quote}
Formalizing, in a precise sense, the connection between reversible programs
based on combinators and paths in HoTT, as intuitive as it may seem, is however
difficult. Paths in HoTT come equipped with principles like the
``contractibility of singletons'', ``transport'', and ``path induction.'' None
of these principles seems to have any direct counterpart in the world of
reversible programming.

We will however show, in a very clear sense, how a $\Pi$-based language is
``equivalent'' to a univalent HoTT universe that possesses all the required
properties of paths. The essence of the construction can be demonstrated with a
tiny ``featherweight'' variant of $\Pi$ that includes just one type: booleans,
and a tiny ``featherweight'' univalent universe to be formalized in the next
section. Here we give a formalization of the small $\Pi$-based language which we
will use in the formalization in the rest of the paper.

The restriction of $\Pi$ to the case of just booleans is the following:

\[\begin{array}{rcl}
\tau &::=& \bt \\
\\
v &::=& \begin{array}[t]{lrcl}
                    & \fc &:& \bt \\
              \alt & \tc &:& \bt
               \end{array} \\
\\
c &::=& \begin{array}[t]{llcl}
              & \id &:& \tau \iso \tau \\
               \alt & \swap &:& \bt \iso \bt \\
               \alt & ! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
               \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3) \to (\tau_1 \iso \tau_3)
               \end{array} \\
\\
\alpha &::=& \begin{array}[t]{lrcl}
               & \id &:& c \isotwo c \\
            \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
            \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
            \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
            \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
            \alt & \rho &:& \swap \circ \swap \isotwo \id \\
            \alt & \assocc &:& \compc{(\compc{c_1}{c_2})}{c_3} \isotwo \compc{c_1}{(\compc{c_2}{c_3})} \\
            \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2') \to (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
            \alt & !! &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_1) \\
            \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3) \to (c_1 \isotwo c_3)
             \end{array}
\end{array}\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}

Following the HoTT book and the slides on ``A Characterization of Univalent
Fibrations'' by Dan Christensen. All the definitions and most of the examples in
this section are formalized in the accompanying Agda code.

%%%%%
\subsection{Definitions}

\paragraph*{Type Family.} A dependent function $P : A \to \mathcal{U}$ defines a
\emph{type family}, $P(x)$ indexed by $x:A$.

\paragraph*{Functors (Lemma 2.2.1 in the book)} Suppose $f : A \to B$ is a
function. Then for any $x,y:A$ there is an operation
$\mathit{ap}_f : (x \equiv y) \to (f(x) \equiv f(y))$.

\paragraph*{Equivalences (Chapter 4 in the book)} For any two types $A$ and $B$
we can form the type $A \simeq B = \Sigma_{(f : A \to B)} \mathit{isequiv}~f$ of
equivalences. Each equivalence consists of a function $f : A \to B$ together
with evidence $\mathit{equiv}~f$ that this function is an equivalence.

\paragraph*{Paths to Equivalences (Lemma 2.10.1 in the book)} Given two types
$A$ and $B$, there is a function
$\mathit{idtoeqv} : (A \equiv B) \to (A \simeq B)$ mapping paths to
equivalences.

\paragraph*{Transport (Lemma 2.3.1 in the book)} Given a type family
$P : A \to \mathcal{U}$, a path $p : x \equiv y$ in $A$ induces a function
$p_* : P(x) \to P(y)$ which transports points in $P(x)$ to points in $P(y)$.

\paragraph*{Fibration (Sections 2.3 and 2.7 in the book)} We think of a type
family $P : A \to \mathcal{U}$ as a \emph{fibration} with base space $A$, with
$P(x)$ being the fiber over $x$, and with $\Sigma_{(x:A)} P(x)$ being the
\emph{total space} of the fibration, and with first projection
$\Sigma_{(x:A)} P(x) \to A$. In the figure below, $p_*(u)$ is the transport of
$u$ along $p$ and both pairs $(x,u)$ and $(y,p_*(y))$ are in the total space. A
critical property of fibrations is that a path $p : x \equiv y$ in the base
space $A$ and a point $u$ in the fiber over $x$ induce a function that
\emph{lifts} the path to the total space: the lifted path starts at $(x,u)$ and
ends at $(y,p_*(u))$. A function $f : \Pi_{(x:A)} P(x)$ is sometimes called a
\emph{section} of the fibration $P$.

\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (-3,0) ellipse (1.5cm and 3cm);
  \draw (3,2) ellipse (0.8cm and 2cm);
  \draw (3,-2) ellipse (0.5cm and 1cm);
  \node[blue,ultra thick,above] at (-3,3) {$A$};
  \node[blue,ultra thick,above] at (3,3.9) {$P(x)$ (fiber over $x$)};
  \node[blue,ultra thick,below] at (3,-3) {$P(y)$ (fiber over $y$)};
  \draw[fill] (-3,1.5) circle [radius=0.025];
  \draw[fill] (-3,-1.5) circle [radius=0.025];
  \draw[left,cyan,thick] (-3,1.5) -- (-3,-1.5);
  \node[left] (X) at (-3,1.5) {$x$};
  \path[cyan,thick] (X) edge [loop above, looseness=8, in=40, out=140] node[above] {$q$} (X);
  \node[left] at (-3,-1.5) {$y$};
  \draw[fill] (3,-1.8) circle [radius=0.025];
  \draw[fill] (3,3) circle [radius=0.025];
  \node[above] at (3,3) {$u$};
  \draw[fill] (3,2.2) circle [radius=0.025];
  \node[below] at (3,2.2) {$q_*(u)$};
  \draw[fill] (3,1) circle [radius=0.025];
  \node[below] at (3,1) {$v$};
  \node[below] at (3,-1.8) {$p_*(u)$};
  \node[left,cyan] at (-3,0) {$p$};
  \draw[->,red,dashed,ultra thick] (-3,1.5) to [out=45, in=135] (2.2,2.5);
  \draw[->,red,dashed,ultra thick] (-3,-1.5) to [out=-45, in=-135] (2.5,-2.5);
\end{tikzpicture}
\end{center}

\noindent It is worth repeating Remark 2.7.1 from the book. In the figure above,
we could have $(x,u) \equiv (x,v)$ without necessarily having $u \equiv v$. The
only thing we can conclude regarding $u$ and $v$ is that there is a (possibly
non-trivial) path $q : x \equiv x$ such that $q_*(u) \equiv v$.

\paragraph*{Propositional Truncation.} Given a type $A : \mathcal{U}$, the type
$\|A\|$ is the propositional truncation of $A$ defined as follows: for
  any $x:A$ we have $|x| : \|A\|$, and for any $x,y : \|A\|$, we have
  $\gluep : x \equiv y$.

\paragraph*{Univalent Type Family.}
For each fibration, we have a function:
\[\begin{array}{rcl}
\textit {transport-equiv} &:& \Pi_{(P : A \to \mathcal{U})}~ \Pi_{(x,y:A)}~
    x \equiv y  \to P(x) \simeq P(y) \\
\textit{transport-equiv}~P~x~y &=& \lambda p. \mathit{idtoeqv}(\mathit{ap}_{P}(p))
\end{array}\]
We say a type family $P : A \to \mathcal{U}$ is \emph{univalent} if for all
$x,y:A$ the function $\textit{transport-equiv}~P~x~y$ is an
equivalence. Formally:
\[\begin{array}{l}
\textit{is-univalent-fibration} : (A \to \mathcal{U}) \to \mathcal{U} \\
\textit{is-univalent-fibration}~P = \Pi_{(x,y:A)} \textit{isequiv}~(\textit{transport-equiv}~P~x~y)
\end{array}\]

%%%%%
\subsection{Examples}

We will choose particular base spaces $A$ and fibrations $P : A \to \mathcal{U}$
below and reason about whether they are univalent or not. For each fibration,
the question of whether it is univalent reduces to whether, for all $x,y:A$, the
function $\textit{transport-equiv}~P~x~y$ is an equivalence. In other words, the
question we are asking is whether we have an equivalence
$(x \equiv y) \simeq (P(x) \simeq P(y))$ witnessed by the function
$\textit{transport-equiv}~P~x~y$.

\begin{itemize}

\item Take $A = \mathcal{U}$ and $P = \mathit{Id}_{\mathcal{U}}$. By definition,
  this fibration is univalent if
  $\textit{transport-equiv}~\mathit{Id}_{\mathcal{U}}~x~y = \mathit{idtoeqv}$ of
    type $x \equiv y \to x \simeq y$ is an equivalence. As Axiom 2.10.3 in the
    book states this is exactly the statement of the conventional univalence
    axiom for universe $\mathcal{U}$.

\item Take $A = \ot$ with element $\star$ and $P = \lambda \_. \ot$. This
    fibration $P$ is univalent if we have an equivalence
    $(\star\equiv\star) \simeq (\ot\simeq\ot)$ witnessed by
    $\textit{transport-equiv}~P~x~y$. The two types are clearly equivalent in
    this case.

\item Take $A = \ot$ and $P = \lambda \_. \zt$. This fibration is univalent if
    we have an equivalence $(\star\equiv\star)\simeq(\zt\simeq\zt)$ witnessed by
    $\textit{transport-equiv}~P~x~y$. The two types are also clearly equivalent
    in this case.

\item Take $A = \ot$. For no $P$ other than the above two instances will we
    have a univalent fibration.

\item Take $A = \bt$ with elements $\fc$ and $\tc$ and
  $P = \lambda \_. \ot$. The statement
  $\textit{is-univalent-fibration}~P$ reduces to $(\fc \equiv \tc) \simeq
  (\ot\simeq\ot)$ which is false.

\item Take $A = \bt$ with elements $\fc$ and $\tc$ and
  $P = \lambda \{ \fc \to \zt; \tc \to \ot \}$. This fibration is univalent if
  we have an equivalence $(\fc \equiv \tc) \simeq (\zt\simeq\ot)$ which is
  true. That is the only choice of $P$ that works (modulo symmetry).

\item For any $n > 2$, taking $A$ as sum of $n$ copies of $\ot$ can never
  give a univalent fibration.

\item For any type $F$ we define
  $\{F\} = \Sigma_{(A : \mathcal{U})} \| F \simeq A \|$. By a proposition of
  Christensen, any fibration with base space $\{F\}$, i.e., any fibration
  $\{F\} \to \mathcal{U}$, is univalent. In particular, we can take the base
  space to be:
  \begin{itemize}
  \item $\{\ot\}$. The type $\{\ot\}$ is actually the same as $\ot$ and this was
    covered by previous examples.
  \item $\{S^1\}$. The type $\{S^1\}$ is the same as $S^1$. This is perhaps the
    simplest base space to choose but it does not naturally correspond to a
    conventional programming language.
  \item $\{\bt\}$. In some sense this is the simplest
    programming-language--relevant universe --- our featherweight HoTT. We aim
    to characterize this base space as a reversible programming language based
    on type isomorphisms.
  \end{itemize}

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analyzing $\{\bt\}$ and its Connection to $\Pi$}

%%%%%
\subsection{The Subuniverse $\{\bt\}$}

Recall that $\{\bt\} = \Sigma_{(A:\mathcal{U})} \|A \simeq \bt\|$. We begin by
unpacking the objects, 1-paths, and 2-paths of $\{\bt\}$.

\paragraph*{Objects in $\{\bt\}$.} The only term in $\{\bt\}$ is
$(\bt,\|\reflp_{\bt}\|)$ which we call $`\bt$. Formally we can prove
$\|(X,p) \equiv `\bt\|$ is inhabited for all $X$ and $p$.

\paragraph*{1-Paths in $\{\bt\}$.} These are paths between $`\bt$ and
$`\bt$. Generally speaking, paths between elements of a $\Sigma$-type are pairs
of paths with a transport in the second component. If we have a type $A$ with
points $a$ and $b$ and a path $p : a \equiv b$. Fix some $c$ and consider the
dependent function $(x:A) \rightarrow (x \equiv c)$. This forms the types
$a\equiv c$ and $b\equiv c$ with elements $ua$ and $ub$. Now a path between
$(a,ua)$ and $(b,ub)$ is a pair $(p,\alpha)$ where $\alpha : p_* ua \equiv
ub$. Hence paths between $`\bt$ and $`\bt$ are going to be for the form
$(p,\alpha)$ where $p : \bt \equiv \bt$ and $\alpha$ is essentially $\gluep$
from the definition of truncation. So we have two paths:
\[\begin{array}{rcl}
`\mathbf{id} &=& (\reflp,\gluep) \\
`\mathbf{not} &=& (\notp,\gluep)
\end{array}\]

\paragraph*{2-Paths in $\{\bt\}$.} Now are considering paths between
$`\mathbf{id}$ and $`\mathbf{id}$, and between $`\mathbf{not}$ and
$`\mathbf{not}$. We have at least one path at this level which is
$(\reflp,\gluep)$ but to show that that this is the only path we will have to
first prove that $`\mathbf{not} \circ `\mathbf{not} \equiv `\mathbf{id}$.

%%%%%
\subsection{Equivalence}

Now we claim that $\{\bt\}$ is equivalent to $\mathcal{U}_\Pi$ which is the
universe which includes types $\tau$ as a points, combinators
$c : \tau_1 \iso \tau_2$ as 1-paths, and 2-combinators
$\alpha : c_1 \isotwo c_2$ as 2-paths. The proof is in a meta-logic which is
itself univalent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%
\section{Generalization I}

Use all finite types instead of just Bool

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%
\section{Generalization II}

Add a HIT to the univalent universe, perhaps something to do with fractionals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{acm}
{\footnotesize
\bibliography{cites}
}
\end{document}