### 2017-06-30 upi.tex
remove some old material which was entirely subsumed and no longer served any purpose.

\documentclass{entcs}
\usepackage{prentcsmacro}

\def\lastname{Carette, Chen, Choudhury, Rose, and Sabry}

%% Amr
%% words to remember :-)
%% sublime unfathomable
%% path categorical semantics
%% ---

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}
\usepackage[inline]{enumitem}
\usepackage{agda}

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\byisotwo}[1]{{\Leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\textit{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\textit{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}

\newcommand{\Typ}{\mathbf{Type}}
\newcommand{\alt}{~\mid~}
\newcommand{\patht}[1]{\textsc{PATHS}(#1,#1)}
\newcommand{\fpatht}[1]{\textsc{FREEPATHS}(#1,\Box)}
\newcommand{\fpathp}[2]{\textsc{freepath}~#1~#2}
\newcommand{\pathind}[2]{\textsc{pathind}~#1~#2}
\newcommand{\invc}[1]{!\;#1}
\newcommand{\evalone}[2]{eval(#1,#2)}
\newcommand{\evalbone}[2]{evalB(#1,#2)}
\newcommand{\reflp}{\textsc{refl}}
\newcommand{\notp}{\textsc{not}}
\newcommand{\gluep}{\textsc{glue}}
\newcommand{\reflh}{\mathit{refl}_{\sim}}
\newcommand{\symh}[1]{\mathit{sym}_{\sim}~#1}
\newcommand{\transh}[2]{\mathit{trans}_{\sim}~#1~#2}
\newcommand{\reflq}{\mathit{refl}_{\simeq}}
\newcommand{\symq}[1]{\mathit{sym}_{\simeq}~#1}
\newcommand{\transq}[2]{\mathit{trans}_{\simeq}~#1~#2}
\newcommand{\isequiv}[1]{\mathit{isequiv}(#1)}
\newcommand{\idc}{\mathit{id}_{\boolt}}
\newcommand{\swapc}{\mathit{swap}_{\boolt}}
\newcommand{\assocc}{\mathit{assoc}}
\newcommand{\invl}{\mathit{invl}}
\newcommand{\invr}{\mathit{invr}}
\newcommand{\invinv}{\mathit{inv}^2}
\newcommand{\idlc}{\mathit{idl}}
\newcommand{\idrc}{\mathit{idr}}
\newcommand{\swapswap}{\swapc^2}
\newcommand{\compsim}{\compc_{\isotwo}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\isotwo}{\Leftrightarrow}
\newcommand{\piso}{\multimapdotbothB~~}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\bt}{\mathbb{2}}
\newcommand{\fc}{\mathit{false}}
\newcommand{\tc}{\mathit{true}}
\newcommand{\boolt}{\mathbb{B}}
\newcommand{\univ}{\mathcal{U}}
\newcommand{\uzero}{\mathcal{U}_0}
\newcommand{\uone}{\mathcal{U}_1}
\newcommand{\Rule}[2]{
\makebox{
$\displaystyle
\frac{\begin{array}{l}#1\\\end{array}}
{\begin{array}{l}#2\\\end{array}}$}}
\newcommand{\proves}{\vdash}
\newcommand{\jdgg}[3]{#1 \proves #2 : #3}
\newcommand{\jdg}[2]{\proves #1 : #2}
\newcommand{\jdge}[3]{\proves #1 = #2 : #3}
%% codes
%% denotations

\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Amr says: {#1}}\end{minipage}}}
\newcommand{\jacques}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Jacques says: {#1}}\end{minipage}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frontmatter}
\title{From Reversible Programs to \\ Univalent Universes and Back}
\author{Jacques Carette}
\address{McMaster University}
\author{Chao-Hong Chen}
\address{Indiana University}
\author{Vikraman Choudhury}
\author{Robert Rose}
\author{Amr Sabry}
\address{Indiana University}

\begin{abstract}
  We establish a close connection between a reversible programming language
  based on type isomorphisms and a formally presented univalent universe. The
  correspondence relates combinators witnessing type isomorphisms in the
  programming language to paths in the univalent universe; and combinator
  optimizations in the programming language to 2-paths in the univalent
  universe. The result suggests a simple computational interpretation of paths
  and of univalence in terms of familiar programming constructs whenever the
  universe in question is computable.
\end{abstract}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The proceedings of the $2012$ Symposium on Principles of Programming
Languages~\cite{Field:2012:2103656} included two apparently unrelated papers:
\emph{Information Effects} by James and Sabry and \emph{Canonicity for
  2-dimensional type theory} by Licata and Harper. The first paper, motivated by
the physical nature of
computation~\cite{Landauer:1961,PhysRevA.32.3266,Toffoli:1980,bennett1985fundamental,Frank:1999:REC:930275},
proposed, among other results, a reversible language $\Pi$ in which every
program is a type isomorphism. The second paper, motivated by the connections
between homotopy theory and type theory~\cite{vv06,hottbook}, proposed a
judgmental formulation of intensional dependent type theory with a
twice-iterated identity type. During the presentations and ensuing discussions
at the conference, it became apparent, at an intuitive and informal level, that
the two papers had strong similarities. Formalizing the precise connection was
far from obvious, however.

In this paper we report on a formal connection between appropriately formulated
reversible languages on one hand and univalent universes on the other. In the
next section, we give a rational reconstruction of $\Pi$ focusing on a small
``featherweight'' fragment. In Sec.~\ref{sec:univalent}, we review
\emph{univalent fibrations} which allow us to give formal presentations of
``small'' univalent universes. In Sec.~\ref{sec:connection} we state and prove
the formal connection between the systems presented in the preceding two
sections. Sec.~\ref{sec:conclusion} puts our work in a larger context, discusses
related and future work, and concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Simple Reversible Programming Language}

Starting from the physical principle of ``conservation of
information''~\cite{Hey:1999:FCE:304763,fredkin1982conservative}, James and
Sabry~\cite{James:2012:IE:2103656.2103667} propose a family of programming
languages $\Pi$ in which computation preserve information. Technically,
computations are \emph{type isomorphisms} which, at least in the case of finite
types, clearly preserve entropy in the information-theoretic
sense~\cite{James:2012:IE:2103656.2103667}. We will identify a ``featherweight''
version of $\Pi$ to use in our formal development.  But first we illustrate the
general flavor of the family of languages with some examples.

%%%%%
\subsection{Examples}

The examples below assume a representation of the type of booleans
$\bt$ as the disjoint union $\ot \oplus \ot$ with the left injection
representing $\mathsf{false}$ and the right injection representing
$\mathsf{true}$. Given an arbitrary reversible function
$\AgdaFunction{f}$ of type $a \leftrightarrow a$, we can build the
reversible function $\AgdaFunction{controlled}~\AgdaFunction{f}$ that
takes a pair of type $\bt \otimes a$ and checks the incoming boolean;
if it is false (i.e., we are in the left injection), the function
behaves like the identity; otherwise the function applies
$\AgdaFunction{f}$ to the second argument. In either case, the
incoming boolean is reconstituted to maintain reversibility.

{\small
\[\def\arraystretch{1.2}\begin{array}{rcll}
\AgdaFunction{controlled}  &:& \forall a.~ (a \leftrightarrow a) \rightarrow 
                            & (\bt \otimes a \leftrightarrow \bt \otimes a) \\
\AgdaFunction{controlled}~\AgdaFunction{f} &=&

  \bt \otimes a 
    & \byiso{\AgdaFunction{unfoldBool} \otimes \AgdaFunction{id}} \\
&& (\ot \oplus \ot) \otimes a 
    & \byiso{\AgdaFunction{distribute}} \\
&& (\ot \otimes a) \oplus (\ot \otimes a) 
    & \byiso{\AgdaFunction{id} \oplus (\AgdaFunction{id} \otimes \AgdaFunction{f})} \\
&& (\ot \otimes a) \oplus (\ot \otimes a)
    & \byiso{\AgdaFunction{factor}} \\
&& (\ot \oplus \ot) \otimes a
    & \byiso{\AgdaFunction{foldBool} \otimes \AgdaFunction{id}} \\
&& \bt \otimes a & ~ \\
\end{array}
\]}

The $\AgdaFunction{not}$ function is a simple lifting of
\AgdaFunction{swap₊}. Using the
\AgdaFunction{controlled} building block, we can build a
controlled-not $\AgdaFunction{cnot}$ gate and a
controlled-controlled-not gate, also known as the
\AgdaFunction{toffoli} gate. The latter gate is a universal function
for combinational boolean circuits thus showing the expressiveness of
the language.
{\small
\[\begin{array}{rcl}
\AgdaFunction{not} &:& \bt \leftrightarrow \bt \\
\AgdaFunction{not} &=& 
  \AgdaFunction{unfoldBool} \odot \AgdaFunction{swap₊} \odot \AgdaFunction{foldBool} \\
\\ 
\AgdaFunction{cnot} &:& \bt \otimes \bt \leftrightarrow \bt \otimes \bt \\
\AgdaFunction{cnot} &=& \AgdaFunction{controlled}~\AgdaFunction{not} \\
\\ 
\AgdaFunction{toffoli} &:& \bt \otimes (\bt \otimes \bt)
                           \leftrightarrow  \bt \otimes (\bt \otimes \bt) \\
\AgdaFunction{toffoli} &=& \AgdaFunction{controlled}~\AgdaFunction{cnot} \\
\end{array}\]}
%%%

While we wrote \AgdaFunction{controlled} in equation-reasoning style,
\AgdaFunction{not} is written in combinator style.  These are
equivalent as $\byiso{-}$ is defined in terms of $\odot$.

The code for $\AgdaFunction{controlled}~\AgdaFunction{f}$ provides
constructive evidence (i.e., a program, a logic gate, or a hardware
circuit) for an automorphism on $\bt \otimes a$: it can be read
top-down or bottom-up to go back and forth. This perspective naturally
raises the question of when two programs are ``equivalent.'' Consider
the follwing six programs of type~$\bt \leftrightarrow \bt$

{\small 
\[\def\arraystretch{1.2}\begin{array}{rcl}
\AgdaFunction{id₁}~\AgdaFunction{id₂}~\AgdaFunction{id₃}~ 
  \AgdaFunction{not₁}~\AgdaFunction{not₂}~\AgdaFunction{not₃} &:& \bt \leftrightarrow \bt \\
\AgdaFunction{id₁} &=& 
  \AgdaFunction{id} \odot \AgdaFunction{id} \\
\AgdaFunction{id₂} &=& 
  \AgdaFunction{not} \odot \AgdaFunction{id} \odot \AgdaFunction{not} \\
\AgdaFunction{id₃} &=& 
  \AgdaFunction{uniti⋆} \odot \AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{id} \otimes \AgdaFunction{id}) \odot 
                        \AgdaFunction{swap⋆} \odot 
                        \AgdaFunction{unite⋆} \\
\AgdaFunction{not₁} &=& 
  \AgdaFunction{id} \odot \AgdaFunction{not} \\
\AgdaFunction{not₂} &=& 
  \AgdaFunction{not} \odot \AgdaFunction{not} \odot \AgdaFunction{not} \\
\AgdaFunction{not₃} &=& 
  \AgdaFunction{uniti⋆} \odot \AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{not} \otimes \AgdaFunction{id}) \odot 
                        \AgdaFunction{swap⋆} \odot 
                        \AgdaFunction{unite⋆} 
\end{array}\]}

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=0.9,every node/.style={scale=0.9}]
  \draw (1,2) ellipse (0.5cm and 0.5cm);
  \draw[fill] (1,2) circle [radius=0.025];
  \node[below] at (1,2) {*};

  \draw (0,0) ellipse (0.5cm and 1cm);
  \draw[fill] (0,0.5) circle [radius=0.025];
  \node[below] at (0,0.5) {F};
  \draw[fill] (0,-0.5) circle [radius=0.025];
  \node[below] at (0,-0.5) {T};

  \draw     (1,2)    -- (2,2)      ; %% ()
  \draw     (0,0.5)  -- (2,0.5)    ; %% F
  \draw     (0,-0.5) -- (2,-0.5)   ; %% T

  \draw     (2,2)    -- (3,-0.5)   ;
  \draw     (2,0.5)  -- (3,2)      ;
  \draw     (2,-0.5) -- (3,1)      ;

  \draw     (3,2)    -- (3.5,2)    ;
  \draw     (3,1)    -- (3.5,1)    ;
  \draw     (3,-0.5) -- (3.5,-0.5) ;

  \draw     (3.5,2)    -- (4.5,1)    ;
  \draw     (3.5,1)    -- (4.5,2)    ;
  \draw     (3.5,-0.5) -- (4.5,-0.5) ;

  \draw     (4.5,2)    -- (5,2)    ;
  \draw     (4.5,1)    -- (5,1)    ;
  \draw     (4.5,-0.5) -- (5,-0.5) ;

  \draw     (5,2)    -- (6,0.5)  ;
  \draw     (5,1)    -- (6,-0.5) ;
  \draw     (5,-0.5) -- (6,2)    ;

  \draw     (6,2)    -- (7,2)    ;
  \draw     (6,0.5)  -- (8,0.5)  ;
  \draw     (6,-0.5) -- (8,-0.5) ;

  \draw (7,2) ellipse (0.5cm and 0.5cm);
  \draw[fill] (7,2) circle [radius=0.025];
  \node[below] at (7,2) {*};

  \draw (8,0) ellipse (0.5cm and 1cm);
  \draw[fill] (8,0.5) circle [radius=0.025];
  \node[below] at (8,0.5) {F};
  \draw[fill] (8,-0.5) circle [radius=0.025];
  \node[below] at (8,-0.5) {T};

\end{tikzpicture}
\end{center}
\caption{\label{fig:not}Graphical representation of \AgdaFunction{not₃}}
\end{figure}

\noindent We neither expect, nor desire, that all these programs of
the same type to be be equivalent. In the above, each group of
three does represent ``the same'' program but the two groups
represent distinct programs. Thinking extensionally, i.e., by looking
at all possible input-output pairs, it is indeed easy to verify this
fact, but we can do better. We can provide \emph{evidence} (i.e., a
program that manipulates programs) that can constructively transform
every program to an equivalent one. We show such a level-2 program
proving that $\AgdaFunction{not₃}$ is equivalent to
$\AgdaFunction{not}$. For illustration, the program for
$\AgdaFunction{not₃}$ is depicted in Fig.~\ref{fig:not}. We encourage
the reader to map the steps below to manipulations on the diagram that
would incrementally simplify it:

{\small 
\[\def\arraystretch{1.2}\begin{array}{rcll}
\AgdaFunction{notOpt} &:& \AgdaFunction{not₃} \Leftrightarrow \AgdaFunction{not} \\
\AgdaFunction{notOpt} &=& 
  \AgdaFunction{uniti⋆} \odot (\AgdaFunction{swap⋆} \odot 
                        ((\AgdaFunction{not} \otimes \AgdaFunction{id}) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆})))
 & \quad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{assocLeft}} \\
&& \AgdaFunction{uniti⋆} \odot (\AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{not} \otimes \AgdaFunction{id})) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆})
 & \quad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{swapLeft} 
                                  \boxdot \AgdaFunction{id})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot \AgdaFunction{swap⋆}) \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆})
 & \quad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{assocRight}} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot (\AgdaFunction{swap⋆} \odot 
                        (\AgdaFunction{swap⋆} \odot \AgdaFunction{unite⋆})))
 & \quad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot \AgdaFunction{assocLeft})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot ((\AgdaFunction{swap⋆} \odot 
                      \AgdaFunction{swap⋆}) \odot \AgdaFunction{unite⋆}))
 & \quad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot (\AgdaFunction{leftInv} \boxdot \AgdaFunction{id}))} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot (\AgdaFunction{id} \odot \AgdaFunction{unite⋆}))
 & \quad\byisotwo{\AgdaFunction{id} \boxdot (\AgdaFunction{id}
                                  \boxdot \AgdaFunction{idLeft})} \\
&& \AgdaFunction{uniti⋆} \odot ((\AgdaFunction{id} \otimes \AgdaFunction{not})
                      \odot \AgdaFunction{unite⋆})
 & \quad\byisotwo{\AgdaFunction{assocLeft}} \\
&& (\AgdaFunction{uniti⋆} \odot (\AgdaFunction{id} \otimes \AgdaFunction{not})) 
                      \odot \AgdaFunction{unite⋆}
 & \quad\byisotwo{\AgdaFunction{unitiLeft} \boxdot \AgdaFunction{id}} \\
&& (\AgdaFunction{not} \otimes \AgdaFunction{uniti⋆}) \odot \AgdaFunction{unite⋆}
 & \quad\byisotwo{\AgdaFunction{assocRight}} \\
&& \AgdaFunction{not} \otimes (\AgdaFunction{uniti⋆} \odot \AgdaFunction{unite⋆})
 & \quad\byisotwo{\AgdaFunction{id} \boxdot \AgdaFunction{leftInv}} \\
&& \AgdaFunction{not} \otimes \AgdaFunction{id}
 & \quad\byisotwo{\AgdaFunction{idRight}} \\
&& \AgdaFunction{not} 
\end{array}\]}

It is worthwhile mentioning that the above derivation could also be drawn as 
one (large!) commutative diagram in an appropriate category, with each
$\byisotwo{-}$ as a $2$-arrow (and representing a natural isomorphism).
See~\cite{Shulmann} for that interpretation.

%%%%%
\subsection{$\Pi_{\mathbb{2}}$}

Having illustrated the general flavor of the $\Pi$ family of
languages, we present in full detail a small $\Pi$-based language
which we will use in the formalization in the rest of the paper. The
language is the restriction of $\Pi$ to the case of just one
type, the type of booleans:

\jacques{the code above uses $\odot$ for 1-composition, 
$\boxdot$ for parallel 2-composition of $\odot$, while the
code below uses $\circ$ and $\odot$ respectively, which is 
quite confusing.  We should pick one notation.}
\[\def\arraystretch{0.8}\begin{array}{l@{\quad}rclrl}
(\textit{Types}) & \tau &::=& \bt \\
\\
(\textit{Terms}) &  v &::=& \fc &:& \bt \\
              && \alt & \tc &:& \bt \\
\\
 (\textit{1-combinators}) &  c &::=& \id &:& \tau \iso \tau \\
               && \alt & \swap &:& \bt \iso \bt \\
               && \alt & ! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
               && \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3) \to (\tau_1 \iso \tau_3)  \\
\\
(\textit{2-combinators}) & \alpha &::=& \id &:& c \isotwo c \\
            && \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
            && \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
            && \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
            && \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
            && \alt & \rho &:& \swap \circ \swap \isotwo \id \\
            && \alt & \assocc &:& 
                                  \compc{(\compc{c_1}{c_2})}{c_3} \isotwo \compc{c_1}{(\compc{c_2}{c_3})} \\
            && \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2') \to 
                             (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
            && \alt & !! &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_1) \\
            && \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3) \to (c_1 \isotwo c_3)
\end{array}\]

The syntactic category $c$ is that of 1-combinators denoting
reversible programs, type isomorphisms, permutations, or equivalences
depending on one's favorite interpretation. There are two primitive
combinators $\id$ and $\swap$ which are closed under inverses $!$ and
sequential composition $\circ$. The syntactic category $\alpha$ is
that of 2-combinators denoting reversible program transformations,
coherence conditions on type isomorphisms, equivalences between
permutations, or program optimizations depending on one's favorite
interpretation.

It is relatively easy to think of a few sound program transformations
to include in the category $\alpha$ of 2-combinators. The set above
however is \emph{complete}. We also have canonical forms (constructively).

\begin{lemma}[Canonical Forms]
  Given a 1-combinator $c : \tau \leftrightarrow \tau$, we either have
  a normalization $c \Leftrightarrow \id$ or a normalization
  $c \Leftrightarrow \swap$.
\end{lemma}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Univalent Fibrations}
\label{sec:univalent}

We now switch gear and focus on the construction of \emph{univalent
  universes}. In order to state the definition of such universes we
need to introduce two notions: propositional equality
$\AgdaDatatype{≡}$ and equivalence $\AgdaDatatype{≃}$.

Propositional equality, or identity types in intensional dependent
type theory, or path types under the HoTT-interpretation, come equipped
with a single constructor \AgdaInductiveConstructor{refl} and an
induction principle $\AgdaFunction{J}$. A peculiar aspect of
$\AgdaFunction{J}$ is that it does \emph{not} enforce that every
element of an identity type
$\AgdaFunction{x}~\AgdaDatatype{≡}~\AgdaFunction{y}$ is
\AgdaInductiveConstructor{refl}. This ``weakness'' has a beautiful
consequence: instead of identifying all proofs and collapsing the
structure of proofs in the process, it instead permits a rich
combinatorial structure to emerge.

The second notion we need
is that of equivalence. We say that two types or spaces $A$ and $B$
are equivalent if there exist functions $f : A \rightarrow B$ and
$g : B \rightarrow A$ that compose to the identity function%
\footnote{Strictly speaking, this is a \emph{quasi-equivalence}, an
equivalence also requires a certain coherence between the compositions.}.
A notable point is that equivalences do not come with an induction principle.
We say a universe $\mathcal{U}$ is univalent if its space of identities
$\AgdaFunction{A} ≡ \AgdaFunction{B}$ is equivalent to its space of
equivalences $\AgdaFunction{A} ≃ \AgdaFunction{B}$.

\jacques{We should probably be precise with our definition of 
equivalence.}

While univalent universes came first, it was later realised that
the more general univalent type families are more convenient to
work with, and allow the constructions of custom universes. 
We thus need to build up some definitions for what this means.
Note that below we will use ``point'' and ``member of a type'',
``space'' and ``type'', ``path'' and ``equivalence'' as
synonyms.  Furthermore our setting is an intuinistic type
theory (with no classical axioms such a choice or excluded
middle added); this implies that all mappings are
\emph{continuous}.

It is very important to remember that while $A$ and $B$ are spaces,
so are $\AgdaFunction{A} ≡ \AgdaFunction{B}$ and
$\AgdaFunction{A} ≃ \AgdaFunction{B}$.  Thus whatever meaningful
questions we can ask of $A$, we can also ask about
$\AgdaFunction{A} ≡ \AgdaFunction{B}$ and
$\AgdaFunction{A} ≃ \AgdaFunction{B}$.

Fig.~\ref{fig:fib} on the left depicts a type family $P : A \rightarrow \mathcal{U}$
mapping each member $x : A$ to a type $P(x)$. As everything is continuous,
we expect this type family to respect identities. In other words if $x:A$ and
$y:A$ are identified using a path $x \equiv y$ (depicted on the right), it must
be the case that the spaces $P(x)$ and $P(y)$ are equivalent
$P(x) \simeq P(y)$. Although we depict the path ``in'' the Base Space $A$
(and that is indeed the case in classical homotopy theory), it is best to
think of the evidence given by $\equiv$ as being outside $A$ (but in $\mathcal{U}$).

Generally speaking, there is no reason to
expect that the spaces $x \equiv y$ and $P(x) \simeq P(y)$ of
``evidence'' that identify $x$ and $y$, and $P(x)$ and $P(y)$ respectively,
are themselves related, never mind equivalent, for all or even any $x$ and $y$.
In the rare cases in which these ``spaces of evidence'' are equivalent, such a
fibration is a \emph{univalent fibration}.

\begin{figure}
\begin{tabular}{c@{\qquad\qquad}c}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.5,-2) {Fiber $P(x)$};
  \draw (1,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.5,-2) {Fiber $P(y)$};
\end{tikzpicture}
&
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1.3,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.8,-2) {Fiber $P(x)$};
  \draw (1.3,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.8,-2) {Fiber $P(y)$};
  \draw[below,cyan,thick] (-1,-5) -- (1,-5);
  \node[below,cyan,thick] at (0,-5) {$\equiv$};
  \draw[->,red,ultra thick] (-0.8,-1.7) to [out=45, in=135] (0.8,-1.7);
  \draw[->,red,ultra thick] (0.8,-2.3) to [out=-135, in=-45] (-0.8,-2.3);
  \node[red,ultra thick] at (0,-2) {$\simeq$};
\end{tikzpicture}
\end{tabular}
\caption{\label{fig:fib}(left) Type family $P : A \rightarrow \mathcal{U}$ as a
  fibration with total space $\Sigma_{(x:A)} P(x)$;\newline
 (right) a path $x \equiv y$
  in the base space induces an equivalence between the spaces $P(x)$ and $P(y)$}
\end{figure}

As is often the case when confronted with new definitions, it is
instructive to disect \emph{non-examples}, as a means to understand
what ``goes wrong''.  As a first example of a non-univalent fibration,
consider $P : \mathbb{1} \rightarrow \mathbb{U}$ defined as $P =
\lambda\_.\mathbb{2}$.

\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $\mathbb{1}$};
  \draw[fill] (0,-5) circle [radius=0.025];
  \node[below] at (0,-5) {$\star$};
  \draw (-1.3,-2.7) ellipse (0.5cm and 1cm);
  \draw[fill] (-1.5,-2.7) circle [radius=0.025];
  \draw[fill] (-1.1,-2.7) circle [radius=0.025];
  \node[left] at (-1.8,-2.7) {Fiber $P(\star) = \mathbb{2}$};
  \draw (1.3,-2.7) ellipse (0.5cm and 1cm);
  \draw[fill] (1.1,-2.7) circle [radius=0.025];
  \draw[fill] (1.5,-2.7) circle [radius=0.025];
  \node[right] at (1.8,-2.7) {Fiber $P(\star) = \mathbb{2}$};
  \node[below,cyan,dashed,thick] at (0,-5) {$\equiv$};
  \draw[->,red,dashed,ultra thick] (-0.8,-2.3) to [out=45, in=135] (0.8,-2.3);
  \draw[->,red,dashed,ultra thick] (0.8,-3.1) to [out=-135, in=-45] (-0.8,-3.1);
  \node[red,ultra thick] at (0,-2.7) {$\simeq$};
\end{tikzpicture}

\medskip\noindent There is actually \emph{a single fiber}, but
the identity function and boolean negation induce \emph{two distinct
equivalences} of the space $P(\star) = \mathbb{2}$ with itself.
In other words, $\mathbb{2} \simeq \mathbb{2}$ has two distinct inhabitants.
As there is only one path $\star \equiv \star$ in the base space, the
fibration $P$ is not univalent.

Another example is given by $Q : \mathbb{2} \rightarrow \mathbb{U}$ defined as
$Q = \lambda \_. \mathbb{1}$ illustrated below

\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $\mathbb{2}$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$\fc$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$\tc$};
  \draw (0,-2.7) ellipse (0.5cm and 1cm);
  \node[right] at (0.5,-2.7) {Fiber $Q(\fc)$ = Fiber $Q(\tc)$ = $\mathbb{1}$};
  \draw[fill] (0,-2.7) circle [radius=0.025];
\end{tikzpicture}

The situation here is different.  There is one equivalence
$Q(\fc) = \mathbb{1} \simeq \mathbb{1} = Q(\tc)$ but it does not come
from a path (such as $\fc \equiv \tc$, which does not exist).  And that
equivalence does not ``induce'' such a path either.

As the examples illustrate there is a delicate
balance required for the fibration to be univalent; otherwise some distinct
equivalences would be identified (like for $P$) or distinct points in the base
space would be identified (like for $Q$).

On the other hand, for $P : \mathbb{2} \rightarrow \mathbb{U}$
defined as $P(\fc) = \mathbb{0}$ and $P(\tc) = \mathbb{1}$, we do
obtain a univalent fibration.

\bigskip
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $\mathbb{2}$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$\fc$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$\tc$};
  \draw (-1.3,-2) ellipse (0.5cm and 1cm);
  \node[left] at (-1.8,-2) {Fiber $P(\fc) = \mathbb{0}$};
  \draw (1.3,-2) ellipse (0.5cm and 1cm);
  \draw[fill] (1.3,-2) circle [radius=0.025];
  \node[right] at (1.8,-2) {Fiber $P(\tc) = \mathbb{1}$};
  \draw[below,cyan,dashed,thick] (-1,-5) -- (1,-5);
  \node[below,cyan,dashed,thick] at (0,-5) {$\equiv$};
  \draw[->,red,dashed,ultra thick] (-0.8,-1.7) to [out=45, in=135] (0.8,-1.7);
  \draw[->,red,dashed,ultra thick] (0.8,-2.3) to [out=-135, in=-45] (-0.8,-2.3);
  \node[red,ultra thick] at (0,-2) {$\simeq$};
\end{tikzpicture}

\medskip\noindent In this case, the space of equivalences
$\mathbb{0} \simeq \mathbb{1}$ is empty, as is the space of
paths $\fc \equiv \tc$. These two are thus vacuously equivalent.

Of course, we can obtain another (trivial) example from $\mathbb{1}$:
$P : \mathbb{1} \rightarrow \mathbb{U}$, defined as
$P(\star) = \mathbb{1}$. $P$ is univalent as both $\star \equiv \star$ and
$\mathbb{1} \simeq \mathbb{1}$ each have a single inhabitant.

\newpage

%%%
\subsection{A Special Class}

Instead of guessing, there is a special class of univalent fibrations:
\[\begin{array}{rcl}
\{F\} &:& \Sigma_{A:\mathcal{U}} \| F \equiv A \| \rightarrow \mathcal{U} \\
\{F\}(A,|p|) &=& A
\end{array}\]

This construction is always a univalent fibration~\cite{XXX}. We can understand
the basic idea as follows. Take two points $(A,|p|)$ and $(B,|q|)$ in the base
space. We want to compare the space $A \simeq B$ with the space
$(A,|p|) \equiv (B,|q|)$.  In one direction, we have $p : F \equiv A$ and
$q : F \equiv B$ and hence $A \equiv B$ and by univalence $A \simeq B$. In the
other direction \ldots

Also explain base space has higher paths \ldots

\begin{verbatim}
Thm: Let P : Type -> Type such that for all X : Type, P(X)
is a proposition. Then Sigma(X,P) is the base of a univalent
fibration.

Furthermore, we have that *all* univalent fibrations
over Type arise this way. We don't need this fact directly,
but it may help in the explanation. (It's certainly cleaner
than making the connection to classifying bundles in
geometry).

Here are some more examples.

is-prop : Type -> Type has the property that is-prop(X)
is a proposition for all types X. So Sigma(X , is-prop) is
the base of a univalent fibration. In other words, this
is the subuniverse consisting of all and only propositions.

is-set : Type -> Type has the property that is-set(X)
is a proposition for all types X. So Sigma(X , is-set) is
the base of a univalent fibration. In other words, this
is the subuniverse consisting of all and only sets.

is-1-type : Type -> Type has the property that
is-1-type(X) is a proposition for all types X. So
Sigma(X , is-1-type) is the base of a univalent fibration.
In other words, this is the subuniverse consisting of all
and only 1-types. (Externally, these are the 1-groupoids
embedded in inf-groupoids.)

But here's a summary of what I wrote last week, which
I think should suffice for this project. Designating a
collection of types as a univalent universe boils down
to defining a mere predicate on the universe. One approach
to defining such a predicate is to generate a set of
names and an interpretation El (a type family over
these names), and ask that the disjunction "is equal
to the interpretation of a name" be exclusive (i.e.,
a proposition). This last step requires that the set
of names be representatives of distinct normal forms.
Repackage these data --- the collection of types, their
names, and the proof that the predicate on the universe
is a proposition -- as a semantic object admitting
various presentations, which when computable, can be
called a "reversible programming language".

Because "the proof that the predicate on the universe
is a proposition" (and univalence), we naturally get a
family of equivalences arising.  At various levels.
When reflected syntactically, we get multiple
presentations of the same semantic object (where
'same' is really up to things 1 level up).  Because
all we can really talk about are equivalences, when
this syntax is interpreted as a programming language,
it is unavoidably reversible.

Right, the predicate being a proposition suffices
(and is necessary) to define an embedding into a
univalent universe.

(Univalence is used axiomatically, as is the rest of
the type theory, to define the universe. But you're
welcome to interpret ITT + univalence in simplicial
sets and forget about univalence altogether: the
univalent universe is a particular simplicial set
defined in classical set theory, as are the rest
of the subuniverses serving as denotational models.)

By "admitting various presentations", I mean
presentation in the sense of group presentation.
For example, the real functions on R^2 which are
symmetries of the square centered at the origin
can be viewed as the dihedral group on 8 elements:
but we're free to present this group in an infinite
number of ways. These presentations can be taken to
be syntax for a language. If a presentation has a
decidable word problem, and you have the algorithm,
then you could be said to have an operational
semantics, and so a programming language of a sort.
This is what, I take it, you and Amr did with Laplaza's
work on presenting free distributive categories
in the first place. There your semantic object is
the syntactic category of his theory of distributive
categories.

What using HoTT as a metatheory facilitates here is
the construction of models which admit presentations
of infinity-groupoids. (You could do it without HoTT.
But the results would likely be less general, say if
the work was done directly in simplicial sets or
cubical sets.) Using this approach, you have a lot
more options now for cooking up reversible PLs now;
just as many as you have decidable presentations of
inf-groupoids.

The 'exclusive disjunction' aspect of being a
proposition seems to have more content: it's a
sort of decidability / normal form property all
rolled into one.

Yes, I was struck by this too. At least insofar
as we continue with this style of defining predicates
on the universe to obtain the intended interpretation
of the types in a rev PL, it seems that a decidable
enumeration of normal forms of the types will already
be needed. But this doesn't seem like a huge constraint
to me; most of the action is at the level of the paths....

I guess that you assumed in "This last step requires
that the set of names be representatives of distinct
normal forms" that these were also complete?  That we
have a decidable enumeration of all normal forms?

I haven't said anything about an actual presentation
of the inf-groupoid which is the univalent universe
of finite types. You want Pi to do this work. So to
get started, we should define a function from the
BNF grammar corresponding to a syntax for a semiring
to the terms of that universe. As it stands, we
already know that those terms are reflected in Names,
so it would be sufficient to define that function
into Names. Ah, is this what you meant by "complete"?
Yes, we have a name for everything in the extension of
that predicate (by construction).

\end{verbatim}

That space {2} is a higher groupoid. Let's look at its objects; its 1-paths; its
2-paths etc.

Show Agda code

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Correspondence}
\label{sec:connection}

In previous work, on $\Pi$ we noted a possible connection with HoTT:
\begin{quote}
  It is, therefore, at least plausible that a variant of HoTT based exclusively
  on reversible functions that directly correspond to equivalences would have
  better computational properties. Our current result is a step, albeit
  preliminary, in that direction as it only applies to finite
  types~\cite{Carette2016}.
\end{quote}
Formalizing, in a precise sense, the connection between reversible programs
based on combinators and paths in HoTT, as intuitive as it may seem, is however
difficult. Paths in HoTT come equipped with principles like the
``contractibility of singletons'', ``transport'', and ``path induction.'' None
of these principles seems to have any direct counterpart in the world of
reversible programming.

Soundness; completeness; etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion, Related Work, and Conclusion}
\label{sec:conclusion}

\paragraph*{Reversible Languages.}
\noindent The practice of programming languages is replete with \emph{ad hoc}
instances of reversible computations: database transactions, mechanisms for data
provenance, checkpoints, stack and exception traces, logs, backups, rollback
recoveries, version control systems, reverse engineering, software transactional
memories, continuations, backtracking search, and multiple-level ``undo''
features in commercial applications. In the early nineties,
Baker~\cite{Baker:1992:LLL,Baker:1992:NFT} argued for a systematic, first-class,
treatment of reversibility. But intensive research in full-fledged reversible
models of computations and reversible programming languages was only sparked by
the discovery of deep connections between physics and
computation~\cite{Landauer:1961,PhysRevA.32.3266,Toffoli:1980,bennett1985fundamental,Frank:1999:REC:930275},
and by the potential for efficient quantum
computation~\cite{springerlink:10.1007/BF02650179}.

The early developments of reversible programming languages started with a
conventional programming language, e.g., an extended $\lambda$-calculus, and either
\begin{enumerate}
\item extended the language with a history
mechanism~\cite{vanTonder:2004,Kluge:1999:SEMCD,lorenz,danos2004reversible}, or
\item imposed constraints on the control flow constructs to make them
reversible~\cite{Yokoyama:2007:RPL:1244381.1244404}.
\end{enumerate}
More modern approaches recognize that reversible programming languages require
a fresh approach and should be designed from first principles without the
detour via conventional irreversible
languages~\cite{Yokoyama:2008:PRP,Mu:2004:ILRC,abramsky2005structural,DiPierro:2006:RCL:1166042.1166047}.

\paragraph*{The $\Pi$ Family of Languages}
\noindent In previous work, Carette, Bowman, James, and
Sabry~\cite{rc2011,James:2012:IE:2103656.2103667,Carette2016} introduced
the~$\Pi$ family of reversible languages based on type isomorphisms and
commutative semiring identities. The fragment without recursive types is
universal for reversible boolean circuits~\cite{James:2012:IE:2103656.2103667}
and the extension with recursive types and trace
operators~\cite{Hasegawa:1997:RCS:645893.671607} is a Turing-complete reversible
language~\cite{James:2012:IE:2103656.2103667,rc2011}. While at first sight,
$\Pi$ might appear \emph{ad hoc},~\cite{Carette2016} shows that it arises
naturally from an ``extended'' view of the Curry-Howard correspondance: rather
than looking at mere \emph{inhabitation} as the main source of analogy between
logic and computation, \emph{type equivalences} becomes the source of analogy.
This allows one to see an analogy between algebra and reversible computation.
Furthermore, this works at multiple levels: that of $1$-algebra (types form a
semiring under isomorphism), but also $2$-algebra (types and equivalences form a
weak Rig Groupoid).  In other words, by taking ``weak Rig Groupoid'' as the
starting semantics, one naturally gets $\Pi$ as the syntax for the language of
proofs of isomorphisms -- in the same way that many terms of the
$\lambda$-calculus arise from Cartesian Closed Categories.

By restricting to finite types, two operational semantics emerge:
\begin{enumerate}
\item a (reversible) operational semantics of programs over finite types, and
\item a (reversible) operational semantics of program transformers of such programs.
\end{enumerate}
$\Pi$ programs can be interpreted as generalized permutations, as they work on
structure as well as the underlying skeleton of finite types, aka the natural numbers.
Many of the transformers can be \emph{oriented}, and thus be interpreted as
\emph{optimizations}.  As with traditional languages, there are different aspects
we can optimize for (program size, number of steps to completion, etc); we will
not touch on this further in this paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Old}

%%%%%
\subsection{Definitions}

\paragraph*{Type Family.} A dependent function $P : A \to \mathcal{U}$ defines a
\emph{type family}, $P(x)$ indexed by $x:A$.

\paragraph*{Functors (Lemma 2.2.1 in the book)} Suppose $f : A \to B$ is a
function. Then for any $x,y:A$ there is an operation
$\mathit{ap}_f : (x \equiv y) \to (f(x) \equiv f(y))$.

\paragraph*{Equivalences (Chapter 4 in the book)} For any two types $A$ and $B$
we can form the type $A \simeq B = \Sigma_{(f : A \to B)} \mathit{isequiv}~f$ of
equivalences. Each equivalence consists of a function $f : A \to B$ together
with evidence $\mathit{equiv}~f$ that this function is an equivalence.

\paragraph*{Paths to Equivalences (Lemma 2.10.1 in the book)} Given two types
$A$ and $B$, there is a function
$\mathit{idtoeqv} : (A \equiv B) \to (A \simeq B)$ mapping paths to
equivalences.

\paragraph*{Transport (Lemma 2.3.1 in the book)} Given a type family
$P : A \to \mathcal{U}$, a path $p : x \equiv y$ in $A$ induces a function
$p_* : P(x) \to P(y)$ which transports points in $P(x)$ to points in $P(y)$.

\paragraph*{Fibration (Sections 2.3 and 2.7 in the book)} We think of a type
family $P : A \to \mathcal{U}$ as a \emph{fibration} with base space $A$, with
$P(x)$ being the fiber over $x$, and with $\Sigma_{(x:A)} P(x)$ being the
\emph{total space} of the fibration, and with first projection
$\Sigma_{(x:A)} P(x) \to A$. In the figure below, $p_*(u)$ is the transport of
$u$ along $p$ and both pairs $(x,u)$ and $(y,p_*(y))$ are in the total space. A
critical property of fibrations is that a path $p : x \equiv y$ in the base
space $A$ and a point $u$ in the fiber over $x$ induce a function that
\emph{lifts} the path to the total space: the lifted path starts at $(x,u)$ and
ends at $(y,p_*(u))$. A function $f : \Pi_{(x:A)} P(x)$ is sometimes called a
\emph{section} of the fibration $P$.

\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (-3,0) ellipse (1.5cm and 3cm);
  \draw (3,2) ellipse (0.8cm and 2cm);
  \draw (3,-2) ellipse (0.5cm and 1cm);
  \node[blue,ultra thick,above] at (-3,3) {$A$};
  \node[blue,ultra thick,above] at (3,3.9) {$P(x)$ (fiber over $x$)};
  \node[blue,ultra thick,below] at (3,-3) {$P(y)$ (fiber over $y$)};
  \draw[fill] (-3,1.5) circle [radius=0.025];
  \draw[fill] (-3,-1.5) circle [radius=0.025];
  \draw[left,cyan,thick] (-3,1.5) -- (-3,-1.5);
  \node[left] (X) at (-3,1.5) {$x$};
  \path[cyan,thick] (X) edge [loop above, looseness=8, in=40, out=140] node[above] {$q$} (X);
  \node[left] at (-3,-1.5) {$y$};
  \draw[fill] (3,-1.8) circle [radius=0.025];
  \draw[fill] (3,3) circle [radius=0.025];
  \node[above] at (3,3) {$u$};
  \draw[fill] (3,2.2) circle [radius=0.025];
  \node[below] at (3,2.2) {$q_*(u)$};
  \draw[fill] (3,1) circle [radius=0.025];
  \node[below] at (3,1) {$v$};
  \node[below] at (3,-1.8) {$p_*(u)$};
  \node[left,cyan] at (-3,0) {$p$};
  \draw[->,red,dashed,ultra thick] (-3,1.5) to [out=45, in=135] (2.2,2.5);
  \draw[->,red,dashed,ultra thick] (-3,-1.5) to [out=-45, in=-135] (2.5,-2.5);
\end{tikzpicture}
\end{center}

\noindent It is worth repeating Remark 2.7.1 from the book. In the figure above,
we could have $(x,u) \equiv (x,v)$ without necessarily having $u \equiv v$. The
only thing we can conclude regarding $u$ and $v$ is that there is a (possibly
non-trivial) path $q : x \equiv x$ such that $q_*(u) \equiv v$.

\paragraph*{Propositional Truncation.} Given a type $A : \mathcal{U}$, the type
$\|A\|$ is the propositional truncation of $A$ defined as follows: for
  any $x:A$ we have $|x| : \|A\|$, and for any $x,y : \|A\|$, we have
  $\gluep : x \equiv y$.

\paragraph*{Univalent Type Family.}
For each fibration, we have a function:
\[\begin{array}{rcl}
\textit {transport-equiv} &:& \Pi_{(P : A \to \mathcal{U})}~ \Pi_{(x,y:A)}~
    x \equiv y  \to P(x) \simeq P(y) \\
\textit{transport-equiv}~P~x~y &=& \lambda p. \mathit{idtoeqv}(\mathit{ap}_{P}(p))
\end{array}\]
We say a type family $P : A \to \mathcal{U}$ is \emph{univalent} if for all
$x,y:A$ the function $\textit{transport-equiv}~P~x~y$ is an
equivalence. Formally:
\[\begin{array}{l}
\textit{is-univalent-fibration} : (A \to \mathcal{U}) \to \mathcal{U} \\
\textit{is-univalent-fibration}~P = \Pi_{(x,y:A)} \textit{isequiv}~(\textit{transport-equiv}~P~x~y)
\end{array}\]

%%%%%
\subsection{Examples}

We will choose particular base spaces $A$ and fibrations $P : A \to \mathcal{U}$
below and reason about whether they are univalent or not. For each fibration,
the question of whether it is univalent reduces to whether, for all $x,y:A$, the
function $\textit{transport-equiv}~P~x~y$ is an equivalence. In other words, the
question we are asking is whether we have an equivalence
$(x \equiv y) \simeq (P(x) \simeq P(y))$ witnessed by the function
$\textit{transport-equiv}~P~x~y$.

\begin{itemize}

\item Take $A = \mathcal{U}$ and $P = \mathit{Id}_{\mathcal{U}}$. By definition,
  this fibration is univalent if
  $\textit{transport-equiv}~\mathit{Id}_{\mathcal{U}}~x~y = \mathit{idtoeqv}$ of
    type $x \equiv y \to x \simeq y$ is an equivalence. As Axiom 2.10.3 in the
    book states this is exactly the statement of the conventional univalence
    axiom for universe $\mathcal{U}$.

\item Take $A = \ot$ with element $\star$ and $P = \lambda \_. \ot$. This
    fibration $P$ is univalent if we have an equivalence
    $(\star\equiv\star) \simeq (\ot\simeq\ot)$ witnessed by
    $\textit{transport-equiv}~P~x~y$. The two types are clearly equivalent in
    this case.

\item Take $A = \ot$ and $P = \lambda \_. \zt$. This fibration is univalent if
    we have an equivalence $(\star\equiv\star)\simeq(\zt\simeq\zt)$ witnessed by
    $\textit{transport-equiv}~P~x~y$. The two types are also clearly equivalent
    in this case.

\item Take $A = \ot$. For no $P$ other than the above two instances will we
    have a univalent fibration.

\item Take $A = \bt$ with elements $\fc$ and $\tc$ and
  $P = \lambda \_. \ot$. The statement
  $\textit{is-univalent-fibration}~P$ reduces to $(\fc \equiv \tc) \simeq
  (\ot\simeq\ot)$ which is false.

\item Take $A = \bt$ with elements $\fc$ and $\tc$ and
  $P = \lambda \{ \fc \to \zt; \tc \to \ot \}$. This fibration is univalent if
  we have an equivalence $(\fc \equiv \tc) \simeq (\zt\simeq\ot)$ which is
  true. That is the only choice of $P$ that works (modulo symmetry).

\item For any $n > 2$, taking $A$ as sum of $n$ copies of $\ot$ can never
  give a univalent fibration.

\item For any type $F$ we define
  $\{F\} = \Sigma_{(A : \mathcal{U})} \| F \simeq A \|$. By a proposition of
  Christensen, any fibration with base space $\{F\}$, i.e., any fibration
  $\{F\} \to \mathcal{U}$, is univalent. In particular, we can take the base
  space to be:
  \begin{itemize}
  \item $\{\ot\}$. The type $\{\ot\}$ is actually the same as $\ot$ and this was
    covered by previous examples.
  \item $\{S^1\}$. The type $\{S^1\}$ is the same as $S^1$. This is perhaps the
    simplest base space to choose but it does not naturally correspond to a
    conventional programming language.
  \item $\{\bt\}$. In some sense this is the simplest
    programming-language--relevant universe --- our featherweight HoTT. We aim
    to characterize this base space as a reversible programming language based
    on type isomorphisms.
  \end{itemize}

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analyzing $\{\bt\}$ and its Connection to $\Pi$}

%%%%%
\subsection{The Subuniverse $\{\bt\}$}

\textbf{Theorem.} For any type $F$, the type $\Sigma_{A:\mathcal{U}} \| F \equiv A \|$ is the base
of a univalent fibration.

Recall that
$\{\bt\} = \Sigma_{(A:\mathcal{U})} \|A \simeq \bt\| = \Sigma_{(A:\mathcal{U})}
\| \mathit{Bool} \equiv A\|$. We begin by unpacking the objects, 1-paths, and 2-paths of
$\{\bt\}$.

\paragraph*{Objects in $\{\bt\}$.} The only term in $\{\bt\}$ is
$(\bt,\|\reflp_{\bt}\|)$ which we call $`\bt$. Formally we can prove
$\|(X,p) \equiv `\bt\|$ is inhabited for all $X$ and $p$.

\paragraph*{1-Paths in $\{\bt\}$.} These are paths between $`\bt$ and
$`\bt$. Generally speaking, paths between elements of a $\Sigma$-type are pairs
of paths with a transport in the second component. If we have a type $A$ with
points $a$ and $b$ and a path $p : a \equiv b$. Fix some $c$ and consider the
dependent function $(x:A) \rightarrow (x \equiv c)$. This forms the types
$a\equiv c$ and $b\equiv c$ with elements $ua$ and $ub$. Now a path between
$(a,ua)$ and $(b,ub)$ is a pair $(p,\alpha)$ where $\alpha : p_* ua \equiv
ub$. Hence paths between $`\bt$ and $`\bt$ are going to be for the form
$(p,\alpha)$ where $p : \bt \equiv \bt$ and $\alpha$ is essentially $\gluep$
from the definition of truncation. So we have two paths:
\[\begin{array}{rcl}
`\mathbf{id} &=& (\reflp,\gluep) \\
`\mathbf{not} &=& (\notp,\gluep)
\end{array}\]

\paragraph*{2-Paths in $\{\bt\}$.} Now are considering paths between
$`\mathbf{id}$ and $`\mathbf{id}$, and between $`\mathbf{not}$ and
$`\mathbf{not}$. We have at least one path at this level which is
$(\reflp,\gluep)$ but to show that that this is the only path we will have to
first prove that $`\mathbf{not} \circ `\mathbf{not} \equiv `\mathbf{id}$.

%%%%%
\subsection{Equivalence}

Now we claim that $\{\bt\}$ is equivalent to $\mathcal{U}_\Pi$ which is the
universe which includes types $\tau$ as a points, combinators
$c : \tau_1 \iso \tau_2$ as 1-paths, and 2-combinators
$\alpha : c_1 \isotwo c_2$ as 2-paths. The proof is in a meta-logic which is
itself univalent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%
\section{Generalization I}

Use all finite types instead of just Bool

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%
\section{Generalization II}

Add a HIT to the univalent universe, perhaps something to do with fractionals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{acm}
{\footnotesize
\bibliography{cites}
}
\end{document}