### 2017-10-13 PiFin/paper/main.tex
Start a document to outline what we have

\documentclass{article}

\usepackage{parskip}

\usepackage{macros}

\title{Reversible Programming with Finite Types}
\author{}

\begin{document}

\maketitle

We have the following type families:

\begin{enumerate}
  \item $Fin : \Nat \to \U$
  \item $El : \Nat \to \U$
  \item $CPerm : \Nat \to \Nat \to \U$
\end{enumerate}

We have the following equivalences:

\begin{gather*}
  \equiv{Fin(n)}{El(n)} \\
  \equiv{(Fin(n) \to A)}{List(A, n)} \\
  \equiv{CPerm(n,m)}({\equiv{El(n)}{El(m)}}) \text{ (unproven)} \\
\end{gather*}

We can characterize the path spaces of:

\begin{gather*}
  \Id{\U}{Fin(n)}{Fin(m)} \to \Id{\Nat}{m}{n}
\end{gather*}

We need to show that the subset of Tarski's high school algebra identities
(without exponentials), that is, in the signature $(\Nat, +, \times)$ is
complete.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: