### 2021-05-26 Pi+/paper/reversible.tex

\section{Reversible Programming Languages}~\label{sec:reversible}

\todo{Not the right title.}

\note{This section should explain the main technical parts of the paper
  informally, without using any technology. Use an example, such as, a
  reversible language with $\leq 5$ bits, and examples of permutations and
  transpositions, and when they're equal.}

\note{This might be followed by a section which explains the syntax of Pi.}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End: