### 2021-07-01 Pi+/paper/ufin.tex

\section{The (Sub)Universe of Finite Types}
\label{sec:ufin}

In the previous section, we gave a sound denotational semantics for $\PiLang$ using the groupoid of finite sets and
bijections $\BFin$. Types are interpreted as objects in $\BFin$, and 1-combinators are interpreted as isomorphisms in
$\BFin$. We also showed that if there is a 2-combinator between two 1-combinators, then their interpretations are
extensionally equal in $\BFin$.

However, we want to go a step further, and claim completeness for this semantics, that is, we want to quote back from
the semantics to the syntax. Given a morphism in $\BFin$, we want to produce a 1-combinator in $\PiLang$, and given two
equal morphisms in $\BFin$, we want to produce a 2-combinator witnessing the equality of the quoted 1-combinators.

The groupoid $\BFin$ that we presented, is strict, that is, two equal permutations are extensionally identified, without
an explicit witness for the equality. Our first step is to weaken this groupoid $\BFin$ into a weak 2-category. In a
weak category, equational axioms of the category are replaced by higher invertible cells. For example, there are higher
invertible cells for the left and right units of composition and associativity. The higher cells have their own
coherences as well, the 2-cells can be composed with 1-cells by whiskering, and so on.

HoTT provides a rich internal language for working with weak higher groupoids, using the ``Types are weak
$\infty$-groupoids'' correspondence. We will make this precise by observing that a 1-groupoid in HoTT gives a locally
strict $(2,0)$-category, using the identity type for the higher cells. The groupoid $\BFin$ can be written in HoTT as a
type $\UFin$, which is a 1-groupoid.

Secondly, we would like to like to give a syntactic presentation for this groupoid, by describing exactly how the
1-cells and 2-cells are generated. Note that, in $\BFin$, the set of bijections on a finite set of cardinality $n$ has a
group structure, the identity function is the neutral element, group multiplication is given by function composition,
and the group inverse is given by inverting the bijection. This is the automorphism group of the finite set, which is
also known to be the finite symmetric group $\Sn[n]$.

Groups can be defined by listing the carrier set and the operations of the group, or they can be described using a group
presentation, which gives the generators and relations of a group. The last step of our technical development is to use
a group presentation for $\Sn$ to give a presentation for the higher cells of the groupoid $\UFin$. Using this
characterisation, we will obtain our completeness result.

% In the previous section, we have picked a particular encoding of finite types: as sequences of elements in some
% canonical order.

% This is not a ``free'' presentation. Some properties of the encoding might pollute our semantics. We want to make that
% any theorems / properties of the semantics we prove apply to the source language (under various interpretations).

% This motivates us to use a different semantic representation: a univalent universe of finite types which contains the
% finite types and nothing but the finite types and is defined such as alternative representations of finite types are
% identified. This guarantees that we can get a proper, sound and complete, semantics for our language.

% \begin{center}
% \begin{tikzpicture}[scale=1,every node/.style={scale=1}]
% \usetikzlibrary{arrows}
% \usetikzlibrary{shapes}
% \node[ultra thick, draw=blue, ellipse, minimum width=300pt, minimum height=300pt, align=center] {};
% \node[ultra thick, draw=blue, ellipse, minimum width=70pt, minimum height=40pt, align=center] (a) {$\{ 1 , 2 \} \uplus \{ 3 \}$};
% \node[ultra thick, draw=blue, ellipse, minimum width=70pt, minimum height=40pt, align=center] (b) at ([shift=({70pt,-70pt})]a) {$\{ 1 \} \uplus \{ 2, 3 \}$};
% \node[ultra thick, draw=blue, ellipse, minimum width=70pt, minimum height=40pt, align=center] (c) at
% ([shift=({-70pt,-70pt})]a) {$\{ 10 \} \uplus \{ 20, 30 \}$};
% \node (d) at ([shift=({50pt,-10pt})]a) {$\cdots$};
% \draw[thick,-] (a) to node[draw=none,right] {$\equiv$} (b);
% \draw[thick,-] (b) to node[draw=none,below] {$\equiv$} (c);
% \draw[thick,-] (a) to node[draw=none,left] {$\equiv$} (c);
% \end{tikzpicture}
% \end{center}

\[
  \begin{tikzcd}
    \Fin[3]
    \arrow[""{name=0, anchor=center, inner sep=0}, "{f_{3}}", no head, loop, distance=4em, in=115, out=65]
    \arrow[""{name=0, anchor=center, inner sep=0}, "{f_{2}}", no head, loop, distance=8em, in=125, out=55]
    \arrow[""{name=1, anchor=center, inner sep=0}, "{f_{1}}"', no head, loop, distance=12em, in=135, out=45]
    \arrow["", "{h}", shorten <=3pt, shorten >=3pt, Rightarrow, no head, from=0, to=1]
  \end{tikzcd}
\]

We have $\Fin[3] = \Set{0,1,2} \eqv \unit \sqcup (\unit \sqcup \unit)$ which fixes a particular enumeration of the
elements. Suppose we have a set $X = (\unit \sqcup \unit) \sqcup \unit$, it has the same cardinality as $\Fin[3]$, so it
is represented by the same 0-cell. But, $X$ can be made equivalent to $\Fin[3]$ in many different ways since there are
many bijections between them. One bijection is
$\Set{\inl(\inl(\ttt)) \mapsto 0, \inl(\inr(\ttt)) \mapsto 1, \inr(\ttt) \mapsto 2}$ which can be written in two
different ways by composing more primitive operations, $f_{1} = \assocrp$, or
$f_{2} = \swapp \compc \assoclp \compc \swapp$. Another bijection is
$\Set{\inl(\inl(\ttt)) \mapsto 1, \inl(\inr(\ttt)) \mapsto 2, \inr(\ttt) \mapsto 0}$ which is given by $f_{3} = \ldots$.
Since $f_{1}$ and $f_{2}$ produce the same enumeration of the elements of $X$, they are identified by a homotopy $h$
which is encoded in the 2-cell between them.

At level 0, all we know is that if $X : \UFin[3]$, then X is merely equal to $\Fin[3]$, that is
$\Trunc[-1]{X \id \Fin[3]}$, and we don't have access to the bijection. At level 1, if we know that both $X$ and $Y$ are
\emph{equal} in $\UFin[3]$, then we can extract an equivalence between them, that is, $(X \id Y) \to (X \eqv Y)$.
$\UFin[3]$ being a univalent subuniverse asserts that there are as many elements (upto higher homotopy) in $X \id Y$ as
there are $X \eqv Y$.

Given an operational model of how a type A reduces to a type B
Given a denotational model of which types A and B are equivalent
Want soundness and completeness results
Guarantee that the two notions of equality are themselves “equivalent”
Generalized “identity of indiscernibles”

Previous semantics adequate for reasoning about evaluation and for checking the equivalence of two circuits.

Want NBE, want rewriting system of program equivalence, what else? So we need to move to proof-relevant higher-category
groupoid semantics instead of set-theoretic.

Now the question is, what does it mean for a compiler?

My opinion is that it is still the same, what rewrites to apply depends
on the choices made by the compiler writer, but we can think about that.

%%%
\subsection{Univalent Subuniverses}
\label{sec:univalent}

In order to define the semantics, we seek a mathematical structure satisfying the following properties: (i) it contains structures corresponding to all the finite types and nothing but the finite types, and (ii) it is robust enough to ensure that equivalent encodings of finite types are identified.

In this section, we introduce some basic concepts and notation that we use in Homotopy Type Theory. Then, we define
univalent subuniverses and discuss some specific examples.

\subsection{The Type Theory}~\label{subsec:type-theory}

We work in Homotopy Type Theory (HoTT), and use the language of the HoTT
book~\cite{univalentfoundationsprogramHomotopyTypeTheory2013}, that is, we use intensional Martin-L\"{o}f Type Theory,
with a univalent universe $\UU$, and we use Higher Inductive Types (HITs) for propositional truncations and set
quotients. All arguments will hold in a cubical type theory as well, such
as~\cite*{cohenCubicalTypeTheory2018,angiuliComputationalSemanticsCartesianCubical2019}. We recall a few basic facts to
familiarise the reader with the notation we use. For more details about basics of HoTT, we refer the reader to the HoTT
book~\cite{univalentfoundationsprogramHomotopyTypeTheory2013}.

Our denotational semantics is given using groupoids, and HoTT provides a rich internal language for describing (weak
$\infty$-)groupoids. By formalising our results using general tools and principles from HoTT, we're able to use a proof
assistant to computer-check our proofs. Our results are formalised in the Agda proof assistant using the HoTT-Agda
library, though it could also be formalised using Cubical Agda~\cite{vezzosiCubicalAgdaDependently2019}. \review{Using
  the formalisation of the semantics, we are also able to extract a computable procedure for performing normalisation of
  circuits.}

\subsubsection{Identity Types}

\vc{I'm explaining everything starting from the categorical language. We do care more about the groupoid structure of
  types, and how to encode groupoids using types, since that's what we use in~\cref{sec:finite}.}

In HoTT, the identity type or the equality type is the type of paths or equalities between two terms of the same type.
Given two terms $x:A$ and $y:A$, we write $x \id_{A} y$, or simply $x \id y$, for the equality type between them. In
book HoTT, the identity type is generated by reflexivity $\refl_{x} : x \id_{A} x$, and the eliminator for the identity
type is given by path induction or the $J$-rule. The identity type equips each type with the structure of a weak
$\infty$-groupoid, or a homotopy type. For example, given an arbitrary type (or groupoid) $A$, we list some groupoid
laws that are provable using path induction.

\begin{gather*}
  \begin{aligned}
    \term{sym}   & : (x \id_{A} y) \to (y \id_{A} x)                   \\
    \term{\comp} & : (x \id_{A} y) \to (y \id_{A} z) \to (x \id_{A} z)
  \end{aligned}
  \qquad
  \begin{aligned}
    \term{assoc} & : (p : x \id_{A} y)  (q : y \id_{A} z) (r : z \id_{A} w)    \\
                 & \to (p \comp q) \comp r \id p \comp (q \comp r)             \\
    \term{invr}  & : (p : x \id_{A} y) \to p \comp \term{sym}(p) \id \refl_{x}
  \end{aligned}
\end{gather*}

Functions between types are functors between groupoids. Given a function $f : A \to B$, the functorial action
$\term{ap}_{f}(\blank)$ is given by path induction.

\[
  \term{ap}_{f} : \dfun{x,y:A}{x \id_{A} y \to f(x) \id_{B} f(y)}
\]

Type families are functions from a type to the universe, such as $A \to \UU$, which is an $A$-indexed family of
groupoids. The $\term{transport}$ operation gives the functorial action of paths in the indexing type, which is also
defined by path induction. If $P : A \to \UU$ is a type family, then for a path $x \id_{A} y$, we have

\[
  \transport{P} : \dfun{x,y:A}{x \id_{A} y \to P(x) \to P(y)}
\]

A type family can also be seen as a fibration, from the topological viewpoint. For a type family $P : A \to \UU$ and a
point $x : A$, the type $P(x)$ gives the fiber over $x$. The type $\dsum{x:A}{P(x)}$ is the collection of all the fibers
and is called the total space of $P$. The first projection ${\pi_1 : \dsum{x:A}{P(x)} \to A}$ from the total space to
the base space $A$ is then a fibration. The lifting operation lifts paths in the base space to paths in the total space.
Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have

\[
  \term{lift}(u,p) : (x , u) \id_{\dsum{x:A}{P(x)}} (y , \tr{p}{u})
\]

where $\tr{p}{u}$ is shorthand for $\transport{P,p}(u)$.

\begin{center}
  \begin{tikzpicture}[yscale=.5,xscale=2]
    \draw (0,0) arc (-90:170:8ex) node[anchor=south east] {$A$} arc (170:270:8ex);
    \draw (0,6) arc (-90:170:10ex) node[anchor=south east] {$\dsum{x:A}{P(x)}$} arc (170:270:10ex);
    \draw[->] (0,5.8) -- node[auto] {$\fst$} (0,3.2);
    \node[circle,fill,inner sep=1pt,label=left:{$x$}] (b1) at (-.5,1.4) {};
    \node[circle,fill,inner sep=1pt,label=right:{$y$}] (b2) at (.5,1.4) {};
    \draw[decorate,decoration={snake,amplitude=1}] (b1) -- node[auto,swap] {$p$} (b2);
    \node[circle,fill,inner sep=1pt,label=left:{$\pair{x,u}$}] (b1) at (-.5,7.2) {};
    \node[circle,fill,inner sep=1pt,label=right:{$\pair{y,\tr{p}{u}}$}] (b2) at (.5,7.2) {};
    \draw[decorate,decoration={snake,amplitude=1}] (b1) -- node[auto] {$\term{lift}(u,p)$} (b2);
  \end{tikzpicture}
\end{center}

Further, using the groupoid structure of $A$, we can show that $\term{transport}$ lifts paths to equivalences, we define

\[
  \tptEqv{P} : \dfun{x,y:A}{x \id_{A} y \to P(x) \eqv P(y)}
\]

\todo{Diagram.}

\begin{definition}[Univalent Fibration]
  $P$ is a univalent type family (or simply a univalent fibration) if $\tptEqv{P}$ is an equivalence.
\end{definition}

\subsubsection{Univalence}

Voevodsky's \emph{univalence} principle characterises paths in the universe. It says that equivalent types are equal, or
the following function is an equivalence.

\[
  \ua : A \id_{\UU} B \to A \eqv B
\]

\todo{Revise.}

Alternatively, one can say that the identity type family $\term{id} : \UU \to \UU$ is univalent.

\subsubsection{Higher Inductive Types}

\todo{Explain homotopy types, $\hProp$, $\hSet$, etc.}

\vc{These are placeholder definitions, need informal explanations and references to the book.}

\begin{definition}[Propositional Truncation]
  Given a type $A$, the propositional truncation $\Trunc[-1]{A}$, or simply $\Trunc{A}$, is a higher inductive type
  generated by the following constructors,
  \begin{itemize}
    \item an inclusion function $\trunc{\blank} : A \to \Trunc{A}$,
    \item for each $x, y : \Trunc{A}$, a path $\term{trunc}(x,y) : x \id_{\Trunc{A}} y$,
  \end{itemize}
  such that, given any type $B$ with
  \begin{itemize}
    \item a function $g : A \to B$,
    \item for each $x, y : B$, a path $\term{trunc*}(x,y) : x \id_{B} y$,
  \end{itemize}
  there is a unique function $f : \Trunc{A} \to B$ such that,
  \begin{itemize}
    \item $f(\trunc{a}) \equiv g(a)$
    \item for each $x, y : \Trunc{A}$, $\ap{f}{\term{trunc}(x,y)} \id_{B} \term{trunc*}(f(x),f(y))$.
  \end{itemize}
\end{definition}

\begin{definition}[Set Quotient]
  Given a type $A$ which is an $\hSet$, and a relation $R : A \to A \to \hProp$, the set-quotient $\quot{A}{R}$ is the
  higher inductive type generated by
  \begin{itemize}
    \item an inclusion function $q : A \to \quot{A}{R}$,
    \item for each $x, y : A$ such that $R(x,y)$, a path $q(x) \id_{\quot{A}{R}} q(y)$,
    \item a set truncation, for each $x, y : \quot{A}{R}$ and $r, s : x \id_{\quot{A}{R}} y$, we have $r \id s$,
  \end{itemize}
  with \review{an appropriate induction principle.}
\end{definition}

We recall that quotients in HoTT are \emph{effective}, that is, if $R$ is an equivalence relation, we have
$R(x,y) \eqv (q(x) \id_{\quot{A}{R}} q(y))$.

\subsection{Univalent Subuniverses}

\todo{Explain motivation. Starting from a univalent universe which classifies all types, we want to define a subuniverse
  which classifies only certain types, for example, types that satisfy some property that we want.}

\begin{definition}[Universe]
  A universe \`{a} la Tarski is given by the following pieces of data,
  \begin{itemize}
    \item a code $U : \UU$,
    \item a decoding function $\El : U \to \UU$.
  \end{itemize}
  If $\El$ is a univalent fibration, $U$ is a univalent universe.
\end{definition}

\begin{definition}[Subuniverse]
  A subtype is a type family $P : \UU \to \UU$ whose fibers are prop-valued, that is, $\forall x, \isProp{P(x)}$. A
  subuniverse generated by a subtype has $U \defeq \dsum{X:\UU}{P(X)}$ and $\El \defeq \fst$.
\end{definition}

\begin{proposition}[Univalent Subuniverse]
  Subuniverses generated by subtypes are univalent.
\end{proposition}

\begin{proof}
  Suppose $(U, \El) \defeq (\dsum{X:\UU}{P(X)}, \fst)$ is a subuniverse generated by a subtype $P : \UU \to \UU$. For
  any $X, Y : \UU$ such that $\phi : P(X)$ and $\psi : P(Y)$, we want to show that
  $\tptEqv{\fst} : (X,\phi) \id (Y,\psi) \to X \eqv Y$ is an equivalence. We construct
  $X \eqv Y \to (X,\phi) \id (Y,\psi)$ by $\ua$ and using the fact that $P(\blank)$ is a proposition. That it is an
  inverse follows by calculation using the appropriate computation rules.
\end{proof}

\begin{example}[$\BAut$]
  The type of self-equivalences on a type $T$ is defined as $\Aut[T] \defeq T \eqv T$, which is an $\infty$-group. The
  subuniverse of types that are merely equal to $T$ is given by $\BAut[T] \defeq \Sub{T}$. This turns out to be the
  classifying space of $\Aut[T]$, as justified by~\cref{lem:loop-deloop}. We write $T_{o} \defeq (T, \trunc{\refl_{T}})$
  for the image of the inclusion of $T$.
\end{example}

\begin{example}
  The univalent subuniverse of types merely equal to $\Bool$ is given by $\BAut[\Bool]$. This can be used to give the
  denotational semantics of a 2-bit reversible language~\cite{caretteReversibleProgramsUnivalent2018}.
\end{example}

\begin{definition}[$\Fin$]
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by cardinality. It is defined equivalently in
  two different ways,
  \begin{gather*}
    \begin{aligned}
      \Fin[n] & \defeq \dsum{k:\Nat}{k < n}
    \end{aligned}
    \begin{aligned}
       & \Fin[0] \defeq \bot                      \\
       & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a $\hSet$, and we use the definitions interchangeably.
\end{definition}

\begin{example}
  For any $n : \Nat$, we define the univalent subuniverse of types merely equal to $\Fin[n]$ as $\BAut[\Fin[n]]$. This
  is the universe of finite sets with cardinality equal to $n$. We write $F_{n} \defeq (\Fin[n], \trunc{\refl})$ for the
  image of the inclusion of $\Fin[n]$.
\end{example}

\begin{definition}[$\isFin$]
  We say that a type is finite if it is merely equal to $\Fin[n]$ for some $n$.
  \[
    \isFin[X] \defeq \dsum{n:\Nat}{\SubP{X}{\Fin[n]}}
  \]
  Note that the natural number $n$ need not be truncated, as justified below.
\end{definition}

\begin{proposition}
  For any type $X$, $\isFin[X]$ is a proposition.
\end{proposition}

\begin{proof}
  Suppose we have $(n,\phi) : \isFin[X]$ and $(m,\psi) : \isFin[X]$, we need to show that $(n,\phi) \id (m,\psi)$. It is
  enough to show that $n \id m$. Since $\Nat$ is a set, this is a proposition, so we can use the induction principle of
  propositional truncation to eliminate to $n \id m$, applying it on $\phi$ and $\psi$ respectively. This gives us the
  equalities $X \id \Fin[n]$ and $X \id \Fin[m]$, which gives us $\Fin[n] \id \Fin[m]$, from which $n \id m$ follows by
  applying the first projection.
\end{proof}

\begin{example}
  The univalent subuniverse of \emph{all finite types} is given by
  \[
    \UFin \defeq \dsum{X:\UU}{\isFin[X]}.
  \]
  We write $F_{n} \defeq (\Fin[n], n, \trunc{\refl})$, for the image of the inclusion of $\Fin[n]$. We use this as the
  denotational semantics to interpret our reversible programming language $\PiPlusLang$.
\end{example}

\review{We characterise the path space of univalent subuniverses.}

\begin{proposition}
  If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
\end{proposition}

\begin{proof}
  We need to show that the equality type of $\BAut[T]$ is an $n$-type. Assume $X, Y : \BAut[T]$. Since $\BAut[T]$ is a
  univalent subuniverse, we have $(X \id Y) \eqv (\fst(X) \eqv \fst(Y))$. Note that being an $n$-type is a proposition.
  Since $T$ is an $n$-type, and $\fst(X)$ and $\fst(Y)$ are merely equal to $T$, they're also $n$-types. It follows that
  $\fst(X) \eqv \fst(Y)$ is an $n$-type, and hence $X \id Y$ is an $n$-type.
\end{proof}

\begin{proposition}
  For any $T : \UU$, $\BAut[T]$ is 0-connected.
\end{proposition}

\begin{corollary}
  $\UFin$ is a pointed, connected, 1-type, that is, a 1-groupoid, or has h-level 3.
\end{corollary}

\begin{lemma}~\label{lem:loop-deloop}
  \[
    \loopspace[\BAut[T],T_{0}] \eqv \Aut[T]
  \]
\end{lemma}

\begin{proof}
  Since $\BAut[T]$ is a univalent universe, it follows that
  \[
    (T_{0} \id_{\BAut[T]} T_{0}) \eqv (\fst(T_{0}) \eqv \fst(T_{0})) \equiv (T \eqv T) \equiv \Aut[T].
  \]
\end{proof}

\begin{corollary}
  For every $n:\Nat$,
  \[
    \loopspace[\UFin[n],F_{n}] \eqv \Aut[\Fin[n]]
  \]
\end{corollary}

\[
  \begin{tikzcd}
    F_{0}
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
    & F_{1}
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
    & F_{2}
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=8em, in=125, out=55]
    & \ldots
    & F_{n}
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=8em, in=125, out=55]
    \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=12em, in=135, out=45]
    & \ldots
  \end{tikzcd}
\]

\subsection{Rig structure}~\label{subsec:rig}

\todo{rig}

We describe the symmetric monoidal structure of the groupoid $\UFin$.

First, we observe a few equivalences.

\begin{proposition}
  For any $n, m : \Nat$,
  \begin{align*}
    \Fin[0]                & \eqv \bot        \\
    \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m] \\
  \end{align*}
  and for any types $X, Y, Z$,
  \begin{align*}
    \bot \sqcup X         & \eqv X                     \\
    X \sqcup \bot         & \eqv X                     \\
    (X \sqcup Y) \sqcup Z & \eqv X \sqcup (Y \sqcup Z) \\
    X \sqcup Y            & \eqv Y \sqcup Y            \\
  \end{align*}
\end{proposition}

\begin{proposition}
  For any $n, m : \Nat$,
  \begin{align*}
    \Fin[1]                & \eqv \top        \\
    \Fin[n] \times \Fin[m] & \eqv \Fin[n * m] \\
  \end{align*}
  and for any types $X, Y, Z$,
  \begin{align*}
    \top \times X         & \eqv X                                \\
    X \times \top         & \eqv X                                \\
    (X \times Y) \times Z & \eqv X \times (Y \times Z)            \\
    X \times Y            & \eqv Y \times X                       \\
    X \times \bot         & \eqv \bot                             \\
    X \times (Y \sqcup Z) & \eqv (X \times Y) \sqcup (X \times Z)
  \end{align*}
\end{proposition}

We lift these equivalences to $\UFin$ giving it the additive and multiplicative symmetric monoidal structures,
$(O, \oplus)$ and $(I, \otimes)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$, $\alpha_{X,Y,Z}$,
and the symmetry isomorphism $\mathcal{B}_{X,Y}$. Note that types in $\UFin$ are $\hSet$s since they're merely
equivalent to $\Fin[n]$ for some $n : \Nat$.

\todo{combine both symmetric monoidal structures}

\begin{definition}
  \begin{align*}
    I                 & \defeq F_{0}                                       \\
    X \oplus Y        & \defeq X \sqcup Y                                  \\
    \lambda_{X}       & : I \oplus X \eqv X                                \\
    \rho_{X}          & : X \oplus I \eqv X                                \\
    \alpha_{X,Y,Z}    & : (X \oplus Y) \oplus Z \eqv X \oplus (Y \oplus Z) \\
    \mathcal{B}_{X,Y} & : X \oplus Y \eqv Y \oplus X
  \end{align*}
\end{definition}

These isomorphisms satisfy the Mac Lane coherence conditions for symmetric monoidal categories, that is, the triangle,
pentagon, and hexagon identities, and the syllepsis of the braiding, upto 2-paths in $\UFin$.

\begin{proposition}
  % https://q.uiver.app/?q=WzAsNCxbMCwwLCIoWCBcXG9wbHVzIEkpIFxcb3BsdXMgWSJdLFsyLDAsIlggXFxvcGx1cyAoSSBcXG9wbHVzIFkpIl0sWzEsMSwiWCBcXG9wbHVzIFkiXSxbMCwxXSxbMCwxLCJcXGFscGhhX3tYLEksWX0iXSxbMCwyLCJcXHJob197WH0gXFxvcGx1cyAxX3tZfSIsMl0sWzEsMiwiMV97WH0gXFxvcGx1cyBcXGxhbWJkYV97WX0iXSxbNSw2LCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
  \[\begin{tikzcd}
      {(X \oplus I) \oplus Y} && {X \oplus (I \oplus Y)} \\
      {} & {X \oplus Y}
      \arrow["{\alpha_{X,I,Y}}", from=1-1, to=1-3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{\rho_{X} \oplus 1_{Y}}"', from=1-1, to=2-2]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{1_{X} \oplus \lambda_{Y}}", from=1-3, to=2-2]
      \arrow["\id", Rightarrow, draw=none, from=0, to=1]
    \end{tikzcd}\]
  % https://q.uiver.app/?q=WzAsNSxbMCwxLCIoKFcgXFxvcGx1cyBYKSBcXG9wbHVzIFkpIFxcb3BsdXMgWiJdLFsxLDAsIihXIFxcb3BsdXMgWCkgXFxvcGx1cyAoWSBcXG9wbHVzIFopIl0sWzIsMSwiVyBcXG9wbHVzIChYIFxcb3BsdXMgKFkgXFxvcGx1cyBaKSkiXSxbMiwzLCJXIFxcb3BsdXMgKChYIFxcb3BsdXMgWSkgXFxvcGx1cyBaKSJdLFswLDMsIihXIFxcb3BsdXMgKFggXFxvcGx1cyBZKSkgXFxvcGx1cyBaIl0sWzAsMSwiXFxhbHBoYV97VyBcXG9wbHVzIFgsIFksIFp9Il0sWzEsMiwiXFxhbHBoYV97VyxYLFkgXFxvcGx1cyBafSJdLFszLDIsIjFfe1d9IFxcb3BsdXMgXFxhbHBoYV97WCxZLFp9IiwyXSxbMCw0LCJcXGFscGhhX3tXLFgsWX0gXFxvcGx1cyAxX3tafSIsMl0sWzQsMywiXFxhbHBoYV97VyxYIFxcb3BsdXMgWSxafSIsMl0sWzAsMiwiXFxpZCIsMSx7Im9mZnNldCI6NSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
  \[\begin{tikzcd}
      & {(W \oplus X) \oplus (Y \oplus Z)} \\
      {((W \oplus X) \oplus Y) \oplus Z} && {W \oplus (X \oplus (Y \oplus Z))} \\
      \\
      {(W \oplus (X \oplus Y)) \oplus Z} && {W \oplus ((X \oplus Y) \oplus Z)}
      \arrow["{\alpha_{W \oplus X, Y, Z}}", from=2-1, to=1-2]
      \arrow["{\alpha_{W,X,Y \oplus Z}}", from=1-2, to=2-3]
      \arrow["{1_{W} \oplus \alpha_{X,Y,Z}}"', from=4-3, to=2-3]
      \arrow["{\alpha_{W,X,Y} \oplus 1_{Z}}"', from=2-1, to=4-1]
      \arrow["{\alpha_{W,X \oplus Y,Z}}"', from=4-1, to=4-3]
      \arrow["\id"{description}, shift right=5, draw=none, from=2-1, to=2-3]
    \end{tikzcd}\]
  % https://q.uiver.app/?q=WzAsNixbMSwwLCJYIFxcb3BsdXMgKFkgXFxvcGx1cyBaKSJdLFswLDEsIihYIFxcb3BsdXMgWSkgXFxvcGx1cyBaIl0sWzAsMiwiKFkgXFxvcGx1cyBYKSBcXG9wbHVzIFoiXSxbMSwzLCJZIFxcb3BsdXMgKFggXFxvcGx1cyBaKSJdLFsyLDIsIlkgXFxvcGx1cyAoWiBcXG9wbHVzIFgpIl0sWzIsMSwiKFkgXFxvcGx1cyBaKSBcXG9wbHVzIFgiXSxbMSwwLCJcXGFscGhhX3tYLFksWn0iXSxbMSwyLCJcXG1hdGhjYWx7Qn1fe1gsWX0gXFxvcGx1cyAxX3tafSIsMl0sWzIsMywiXFxhbHBoYV97WSxYLFp9IiwyXSxbMyw0LCIxX3tZfSBcXG9wbHVzIFxcbWF0aGNhbHtCfV97WCxafSIsMl0sWzUsNCwiXFxhbHBoYV97WSxaLFh9Il0sWzAsNSwiXFxtYXRoY2Fse0J9X3tYLFkgXFxvcGx1cyBafSJdLFs3LDEwLCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
  \[\begin{tikzcd}
      & {X \oplus (Y \oplus Z)} \\
      {(X \oplus Y) \oplus Z} && {(Y \oplus Z) \oplus X} \\
      {(Y \oplus X) \oplus Z} && {Y \oplus (Z \oplus X)} \\
      & {Y \oplus (X \oplus Z)}
      \arrow["{\alpha_{X,Y,Z}}", from=2-1, to=1-2]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{\mathcal{B}_{X,Y} \oplus 1_{Z}}"', from=2-1, to=3-1]
      \arrow["{\alpha_{Y,X,Z}}"', from=3-1, to=4-2]
      \arrow["{1_{Y} \oplus \mathcal{B}_{X,Z}}"', from=4-2, to=3-3]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{\alpha_{Y,Z,X}}", from=2-3, to=3-3]
      \arrow["{\mathcal{B}_{X,Y \oplus Z}}", from=1-2, to=2-3]
      \arrow["\id", Rightarrow, draw=none, from=0, to=1]
    \end{tikzcd}\]
  % https://q.uiver.app/?q=WzAsMyxbMCwwLCJYIFxcb3BsdXMgWSJdLFsyLDAsIlggXFxvcGx1cyBZIl0sWzEsMSwiWSBcXG9wbHVzIFgiXSxbMCwxLCIxX3tYIFxcb3BsdXMgWX0iLDAseyJsZXZlbCI6Miwic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFswLDIsIlxcbWF0aGNhbHtCfV97WCxZfSIsMl0sWzIsMSwiXFxtYXRoY2Fse0J9X3tZLFh9IiwyXSxbNCw1LCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
  \[\begin{tikzcd}
      {X \oplus Y} && {X \oplus Y} \\
      & {Y \oplus X}
      \arrow["{1_{X \oplus Y}}", Rightarrow, no head, from=1-1, to=1-3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{\mathcal{B}_{X,Y}}"', from=1-1, to=2-2]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{\mathcal{B}_{Y,X}}"', from=2-2, to=1-3]
      \arrow["\id", Rightarrow, draw=none, from=0, to=1]
    \end{tikzcd}\]
\end{proposition}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End: