### 2021-07-06 Pi+/paper/ufin.tex
unifying notation between sec 2 and sec 3; working on small changes to sec 3

\section{The Groupoid of Finite Types}
\label{sec:ufin}

The denotational semantics for $\PiLang$ in the previous section forms a groupoid $\BFin$ whose objects are sets and
whose (iso)morphisms as bijections. The semantics interprets types as objects in $\BFin$, 1-combinators as isomorphisms
in $\BFin$, and for every pair of 1-combinators related by a 2-combinator, their interpretations are extensionally equal
in $\BFin$. This groupoid is strict: two equal permutations are extensionally identified, without an explicit witness
for the equality. In order to establish completeness, we want to quote back from the semantics to the
syntax. Specifically, given a morphism in $\BFin$, we want to produce a 1-combinator in $\PiLang$, and given two
extensionally equal morphisms in $\BFin$, we want to produce a 2-combinator witnessing the equality of the quoted
1-combinators. Our first step is therefore to \emph{weaken} this groupoid $\BFin$.

In a weak category, equational axioms of the category are replaced by higher invertible cells. For example, there are
higher invertible cells for the left and right units of composition, and associativity. The higher cells come with their
own coherence data as well, such as vertical composition, and horizontal composition, or whiskering.\todo{define
  2-category in appendix.} HoTT provides a rich internal language for working with weak higher groupoids, using the
``Types are weak $\infty$-groupoids'' correspondence. Using the identity type for morphisms, a 1-groupoid in HoTT has
points for 0-cells, 1-paths for 1-cells, and 2-paths for 2-cells, but there is at most one 2-cell between compatible
1-cells.~\footnote{This is a locally-strict $(2,0)$-category, since every cell above 0 is invertible, and the
  hom-categories are posets, that is, truth-value enriched.} The groupoid $\BFin$ can be written in HoTT as a type
$\UFin$, which is a 1-groupoid. In this section, we develop the tools necessary to construct $\UFin$, starting with a
quick review of HoTT.

\begin{toappendix}
  We give an example of the groupoid structure on a 3-element set.

  \[
    \begin{tikzcd}
      \Fin[3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{f_{3}}", no head, loop, distance=4em, in=115, out=65]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{f_{2}}", no head, loop, distance=8em, in=125, out=55]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{f_{1}}"', no head, loop, distance=12em, in=135, out=45]
      \arrow["", "{h}", shorten <=3pt, shorten >=3pt, Rightarrow, no head, from=0, to=1]
    \end{tikzcd}
  \]

  We have $\Fin[3] = \Set{0,1,2} \eqv \unit \sqcup (\unit \sqcup \unit)$ which fixes a particular enumeration of the
  elements. Suppose we have a set $X = (\unit \sqcup \unit) \sqcup \unit$, it has the same cardinality as $\Fin[3]$, so it
  is represented by the same 0-cell. But, $X$ can be made equivalent to $\Fin[3]$ in many different ways since there are
  many bijections between them. One bijection is
  $\Set{\inl(\inl(\ttt)) \mapsto 0, \inl(\inr(\ttt)) \mapsto 1, \inr(\ttt) \mapsto 2}$ which can be written in two
  different ways by composing more primitive operations, $f_{1} = \assocrp$, or
  $f_{2} = \swapp \compc \assoclp \compc \swapp$. Another bijection is
  $\Set{\inl(\inl(\ttt)) \mapsto 1, \inl(\inr(\ttt)) \mapsto 2, \inr(\ttt) \mapsto 0}$ which is given by $f_{3} = \ldots$.
  Since $f_{1}$ and $f_{2}$ produce the same enumeration of the elements of $X$, they are identified by a homotopy $h$
  which is encoded in the 2-cell between them.

  At level 0, all we know is that if $X : \UFin[3]$, then X is merely equal to $\Fin[3]$, that is
  $\Trunc[-1]{X \id \Fin[3]}$, and we don't have access to the bijection. At level 1, if we know that both $X$ and $Y$ are
  \emph{equal} in $\UFin[3]$, then we can extract an equivalence between them, that is, $(X \id Y) \to (X \eqv Y)$.
  $\UFin[3]$ being a univalent subuniverse asserts that there are as many elements (upto higher homotopy) in $X \id Y$ as
  there are $X \eqv Y$.

\end{toappendix}

\begin{toappendix}

  \subsection{To Put Somewhere or Delete}

  This approach is sufficient to prove the semantics forms a 1-category but ignores the rich structure at the next
  level~\cite{carette2016}.
  As explained in the previous section, a $\Pi$-type $A$ has $\sizet{A}$-elements and for all combinators $c : A \iso B$
  we have that $\sizet{A} = \sizet{B}$. Hence, the denotation $\denot{A}$ of a type $A$ with $n$-elements can be the finite
  set $\Fin[n] = \{ 0, 1, \cdots, n-1\}$; the denotation of a value $v : A$ such that $\sizet{A}=n$ will be an index in
  the range $[0,n-1]$, and the denotation $\denot{c}$ of a combinator $c : A \iso B$ such that
  $\sizet{A} = \sizet{B} = n$ will be a function from $\Fin[n]$ to $\Fin[n]$ that permutes the elements. Thus, all types
  with 3 elements will denote $\Fin[3]$ and combinators between them will denote permutations on $\Fin[3]$, e.g.:
  \[\begin{array}{rcl}
      \denot{\onet + (\onet + \onet)}                                         & = & \{ 0,1,2 \} \\
      \denot{(\onet + \onet) + \onet}                                         & = & \{ 0,1,2 \} \\
      \\
      \denot{\assoclp : \onet + (\onet + \onet) \iso (\onet + \onet) + \onet} & = & (0~1~2)     \\
      \denot{\swapp : \onet + (\onet + \onet) \iso (\onet + \onet) + \onet}   & = & (2~0~1)
    \end{array}\]
  where we have used the one-line notation for permutations with $(a~b~c)$ representing the
  permutation that maps 0 to $a$, 1 to $b$, and 2 to $c$. To make the denotation of values precise, we compute a canonical
  enumeration of the elements of each type:
  \[\begin{array}{rcl}
      \mathit{enum}(\zerot)     & = & [~]                                                                                                          \\
      \mathit{enum}(\onet)      & = & [ ~\Acon{tt}~ ]                                                                                              \\
      \mathit{enum}(A + B)      & = & \mathit{map}~\Acon{inj₁}~\mathit{enum}(A) ~\textsf{+\!+}~ \mathit{map}~\Acon{inj₂}~\mathit{enum}(B)          \\
      \mathit{enum}(A \times B) & = & \mathit{concat}~(\mathit{map}~(\lambda v.\mathit{map}~(\lambda w. (v,w))~\mathit{enum}(B))~\mathit{enum}(A))
    \end{array}\]
  \noindent The specification uses a Haskell-like notation for sequences with $\mathit{map}$ as the operation that applies
  a function to each element of a sequence, \textsf{+\!+} as the binary append operation, and $\mathit{concat}$ the
  operation that appends all the subsequences in a sequence.

  Using the definition, we have:
  \[\begin{array}{rcl}
      \mathit{enum}(\onet + (\onet + \onet)) & = & [ \inlv{\Acon{tt}},~\inrv{(\inlv{\Acon{tt}})},~\inrv{(\inrv{\Acon{tt}})} ] \\
      \mathit{enum}((\onet + \onet) + \onet) & = & [ \inlv{(\inlv{\Acon{tt}})},~\inlv{(\inrv{\Acon{tt}})},~\inrv{\Acon{tt}} ]
    \end{array}\]
  Thus, as shown in the diagrams below, $\assoclp~(\inlv{\Acon{tt}})$ applies the permutation $(0~1~2)$ to the index of
  $\inlv{\Acon{tt}}$ which is 0 and produces index 0 in the $(\onet + \onet) + \onet$ type corresponding to value
  $\inlv{(\inlv{\Acon{tt}})}$. Similarly, $\swapp~(\inlv{\Acon{tt}})$ applies the permutation $(2~0~1)$ to the index of
  $\inlv{\Acon{tt}}$ which is 0 and produces index 2 in the $(\onet + \onet) + \onet)$ type corresponding to value
  $\inrv{\Acon{tt}}$.

  \begin{center}
    \input{assoc-perm.tikz}
    \qquad
    \input{swap-perm.tikz}
  \end{center}

  We choose a canonical set of size $n$, called $\mathsf{Fin}~n$, whose elements are natural numbers less than $n$. To
  compute the denotation of a type $A$, we first calculate its size $n = \sizet{A}$. We then construct the canonical set
  $\mathsf{Fin}~n$ and provide the (trivial) evidence that this set is identical to $(\mathsf{Fin}~n)$:

  \[\begin{array}{rcll}
      \sem{A} & = & (\mathsf{Fin}~n, [ n , \mathsf{refl} ]) & \mbox{where}~\sizet{A} = n
    \end{array}\]

  \noindent The denotation $\sem{c}$ of a combinator $c : A \isoone B$ is a path between $\sem{A}$ and $\sem{B}$. If the
  size of $A$ is $m$ and the size of $B$ is $n$, the desired path is between $(\mathsf{Fin}~m, [ m , \mathsf{refl} ])$ and
  $(\mathsf{Fin}~n, [ n , \mathsf{refl} ])$. This path is directly constructed using $\mathit{ap}$ and the fact that $m=n$
  since combinators are always between types of the same size.

  \noindent Finally, given two combinators $p , q : A \isoone B$ and a 2-combinator $\alpha : p \isotwo q$, the denotation
  $\sem{\alpha}$ of $\alpha$ is a path between $\sem{p}$ and $\sem{q}$.

  \note{We use the rig structure of $\UFin$ in~\cref{subsec:rig} to interpret $\PiLang$.}

  We need a formal definition of normal form (canonical form)

  Recalling that the $\lambda$-calculus arises as the internal language of Cartesian Closed Categories
  (Elliott~\cite{Elliott-2017} gives a particularly readable account of this), we can think of $\Pi$ in similar terms, but
  for symmetric Rig Groupoids instead. For example, we can ask what does the equivalence above represent? It is actually a
  ``linear'' representation of a 2-categorial commutative diagram! In fact, it is a painfully verbose version thereof, as
  it includes many \emph{refocusing} steps because our language does not build associativity into its syntax. Categorical
  diagrams usually do.  Thus if we rewrite the example in diagrammatic form, eliding all uses of associativity, but
  keeping explicit uses of identity transformations, we get that \AgdaFunction{swap{-}fl2⇔swap{-}fl1} represents

  \vspace*{3mm}
  \begin{tikzcd}[column sep=normal, row sep=normal]
    && (a+c)+b \arrow [r, "\swapp \oplus\idd", ""{name=U, below}] & (c+a)+b \arrow [dr, "\assocrp"] && \\
    & a+(c+b) \arrow [ur, "\assoclp"] & & & c+(a+b) \arrow [dr, "\idd\oplus\swapp"] &  \\
    a+(b+c) \arrow [ur, "\idd\oplus\swapp"] \arrow [r, "\assoclp"]
    \arrow [dr, "\assoclp"]
    \arrow [ddr, swap, "\assoclp"]
    & (a+b)+c \arrow [r, "\swapp"] &
    c+(a+b) \arrow [r, swap, "\assoclp", ""{name=D, above}]
    & |[alias=Z]| (c+a)+b \arrow [r, "\assocrp"] &c+(a+b) \arrow [r, "\idd\oplus\swapp"] & c+(b+a) \\
    & (a+b)+c \arrow [dr, "\swapp"] &&&& \\
    & (a+b)+c \arrow [dr, swap, "\swapp"] & c+(a+b) \arrow [rr, swap, "\idd", ""{name=DD, above}]
    \arrow [d, Rightarrow, "\idf\, \mathit{idl}\odot{l}"] &&
    c+(a+b) \arrow [ruu, "\idd\oplus\swapp"] & \\
    && c+(a+b) \arrow [rrruuu, bend right = 40, swap, "\idd\oplus\swapp"] && \\
    \arrow[Rightarrow, from=U, to=D, "\mathit{hexagon}\oplus{r}\, \boxdot\, \idf"]
    \arrow[Rightarrow, from=Z, to=DD, swap, "\idf\boxdot\mathit{linv}\odot{l}\,\boxdot\,\idf"]
  \end{tikzcd}

\end{toappendix}

%%%
\subsection{The Type Theory}~\label{subsec:type-theory}
% \subsection{Univalent Subuniverses}
% \label{sec:univalent}

% In order to define the weak groupoid we're after, we seek a mathematical structure satisfying the following properties:
% (i) it contains structures corresponding to all the finite types and nothing but the finite types, and (ii) it is robust
% enough to ensure that equivalent encodings of finite types are identified.

We review some basic concepts and notation from Homotopy Type Theory that we use in the appendix.~\footnote{We use the
language of the HoTT book~\cite{univalentfoundationsprogramHomotopyTypeTheory2013}, that is, we use intensional
Martin-L\"{o}f Type Theory, with a (univalent) universe $\UU$, and a few Higher Inductive Types (HITs) for propositional
truncation and set quotients. All arguments will hold in a Cubical Type Theory as well, such
as~\cite*{cohenCubicalTypeTheory2018,angiuliComputationalSemanticsCartesianCubical2019,vezzosiCubicalAgdaDependently2019}.}
For more details about HoTT, we refer the reader to the HoTT
book~\cite{univalentfoundationsprogramHomotopyTypeTheory2013}.

To recap, types in HoTT have a weak groupoid structure given by the identity type, where points are given by terms of
the type, and higher morphisms are given by the iterated identity type. Functions between types are functors between
groupoids, and type families, or functions to the universe, are indexed families of groupoids. A type family $P : A \to
\UU$ comes equipped with a functor $\pi_1 : \dsum{x:A}{P(x)} \to \UU$ which has a lifting operation giving it the
structure of a fibration (see the appendix for details).

\begin{toappendix}
\subsubsection{Identity Types}

Given two terms $x:A$ and $y:A$, we write $x \id_{A} y$, or simply $x \id y$, for the identity type, which is the type
of equalities or identifications between them. The identity type is generated by reflexivity $\refl_{x} : x \id_{A} x$,
and the eliminator for the identity type is given by path induction or the $J$-rule (\cref{def:path-induction}). This
construction can be iterated, giving the identity type between two terms of an identity type, repeating ad infinitum.
Using the iterated identity type for morphisms, each type is equipped with the structure of a weak $\infty$-groupoid,
where each morphism satisfies groupoid laws only upto a higher one. Given an arbitrary type (or groupoid) $A$, we list
some laws that are provable using path induction.

  \begin{definition}[Path Induction]
    \label{def:path-induction}
    Given a type family $C : \dfun{x,y:A}{(x \id_A y)} \to \UU$, and a function $c : \dfun{x:A}{C(x,x,\refl_x)}$, there is
    a function $f : \dfun{x,y:A}{\dfun{p:x \id_A y}{C(x,y,p)}}$ such that $f(x,x,\refl_x) \defeq c(x)$.
  \end{definition}

\begin{gather*}
  \begin{aligned}
    \term{\inv{\blank}}      & : (x \id_{A} y) \to (y \id_{A} x)                   \\
    \term{\blank\comp\blank} & : (x \id_{A} y) \to (y \id_{A} z) \to (x \id_{A} z)
  \end{aligned}
  \qquad
  \begin{aligned}
    \term{assoc} & : (p : x \id_{A} y)  (q : y \id_{A} z) (r : z \id_{A} w) \\
                 & \to (p \comp q) \comp r \id p \comp (q \comp r)          \\
    \term{invr}  & : (p : x \id_{A} y) \to p \comp \inv{p} \id \refl_{x}
  \end{aligned}
\end{gather*}

A homotopy between functions $f \htpy g$ is given by pointwise equality between them $\dfun{x:A}{f(x) \id_{B} g(x)}$.
The identity type for functions is equivalent to homotopies between them ${(f \id_{A \to B} g)} \eqv {(f \htpy g)}$, by
function extensionality. An equivalence between types $A \eqv B$ is given by a pair of functions between them which
compose to the identity, $f \comp g \htpy \idfunc_{B}$ and $g \comp f \htpy \idfunc_{A}$, and this is equivalent to the
identity type for the universe, $(A \id_{\UU} B) \eqv (A \eqv B)$, by univalence.

Functions between types are functors between groupoids. Given a function $f : A \to B$, there is a functorial action on
the paths given by $\term{ap}$. Type families, that is, types indexed by terms, are simply functions from a type to the
universe, such as $A \to \UU$. For a type family $P : A \to \UU$ and a point $x : A$, the type $P(x)$ is called the
fiber over $x$.  The $\term{transport}$ operation, named $\term{tr}$, lifts paths in the indexing type to functions
between fibers.
%which is an $A$-indexed family of groupoids.

\begin{gather*}
  \term{ap}_{f} : \dfun{x,y:A}{x \id_{A} y \to f(x) \id_{B} f(y)}
  \qquad
  \term{tr}_{P} : \dfun{x,y:A}{x \id_{A} y \to P(x) \to P(y)}
\end{gather*}

The type $\dsum{x:A}{P(x)}$ is the collection of all the fibers and is called the total space of $P$. The first
projection ${\pi_1 : \dsum{x:A}{P(x)} \to A}$ from the total space to the base space $A$ has the structure of a
fibration, that is, there is a lifting operation (\cref{fig:lift}) which lifts paths in the base space to paths in the
total space. Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have

\[
  \term{lift}(u,p) : (x , u) \id_{\dsum{x:A}{P(x)}} (y , \transport{P}{p}{u})
\]

% where $\tr{p}{u}$ is shorthand for $$.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}[yscale=.5,xscale=2]
      \draw (0,0) arc (-90:170:8ex) node[anchor=south east] {$A$} arc (170:270:8ex);
      \draw (0,6) arc (-90:170:10ex) node[anchor=south east] {$\dsum{x:A}{P(x)}$} arc (170:270:10ex);
      \draw[->] (0,5.8) -- node[auto] {$\fst$} (0,3.2);
      \node[circle,fill,inner sep=1pt,label=left:{$x$}] (b1) at (-.5,1.4) {};
      \node[circle,fill,inner sep=1pt,label=right:{$y$}] (b2) at (.5,1.4) {};
      \draw[decorate,decoration={snake,amplitude=1}] (b1) -- node[auto,swap] {$p$} (b2);
      \node[circle,fill,inner sep=1pt,label=left:{$\pair{x,u}$}] (b1) at (-.5,7.2) {};
      \node[circle,fill,inner sep=1pt,label=right:{$\pair{y,\transport{P}{p}{u}}$}] (b2) at (.5,7.2) {};
      \draw[decorate,decoration={snake,amplitude=1}] (b1) -- node[auto] {$\term{lift}(u,p)$} (b2);
    \end{tikzpicture}
  \end{center}
  \caption{Lifting operation in $P$}
  \label{fig:lift}
\end{figure}

\end{toappendix}

\subsection{Univalent Fibrations}

The $\term{transport}$ operation lifts paths in the base space to maps between the fibers. Using the groupoid structure
of $A$, we can show that $\transport{P}{p}{\blank}$ and $\transport{P}{\inv{p}}{\blank}$ form an equivalence -- we name
this map $\tptEqv{P}$ which lifts paths to equivalences of fibers.

\[
  \tptEqv{P} : \dfun{x,y:A}{x \id_{A} y \to P(x) \eqv P(y)}
\]

The type families (or fibrations) we're interested in are the ones where paths in the base space completely determine
the equivalences in the fibers -- these are called univalent fibrations.

\begin{definition}[Univalent Fibration]
  $P$ is a univalent type family (or, $\fst : {\dsum{x:A}{P(x)}} \to A$ is a univalent fibration) if $\tptEqv{P}$ is an
  equivalence.
\end{definition}

The use of the word \emph{univalent} here is a reference to Voevodsky's \emph{univalence} principle. Indeed, univalence
characterises paths in the universe as equivalences between types, which follows from the canonical fibration $\idfunc :
\UU \to \UU$ being univalent.

\begin{toappendix}

\subsubsection{Homotopy Types}

A type is \emph{contractible} (-2-type) if it has a unique element, that is, there is a center of contraction and every
other point is equal to it. A type is a \emph{proposition} (-1-type) if its equality types are contractible, that is, it
has at most one inhabitant. Iterating this, we can define sets or 0-types (whose equality types are propositions) and
1-groupoids or 1-types (whose equality types are sets), and similarly, higher homotopy $n$-types.

\begin{gather*}
  \begin{aligned}
    \isContr{A} & \defeq \dsum{x:A}{\dfun{y:A}{y \id x}} \\
    \isProp{A}  & \defeq \dfun{x,y:A}{\isContr{x \id y}}
  \end{aligned}
  \qquad
  \begin{aligned}
    \isSet{A} & \defeq \dfun{x,y:A}{\isProp{x \id y}} \\
    \isGpd{A} & \defeq \dfun{x,y:A}{\isSet{x \id y}}
  \end{aligned}
\end{gather*}

\subsubsection{Higher Inductive Types}

Higher Inductive Types generalise Inductive Types, by allowing path constructors besides point constructors. While point
constructors generate the elements of the type, path constructors generate equalities between points in the type. We
describe a few basic HITs that we use.

Given a type $A$, the propositional truncation $\Trunc[-1]{A}$, squashes the elements of $A$ turning it into a
proposition. It is given by a HIT with a point constructor $\trunc{\blank} : A \to \Trunc{A}$, and a path constructor
$\term{trunc}(x,y) : x \id_{\Trunc{A}} y$, which equates every pair of points in the truncation
(see~\cref{def:prop-trunc}).

  \begin{definition}[Propositional Truncation]
    \label{def:prop-trunc}
    Given a type $A$, the propositional truncation $\Trunc[-1]{A}$, or simply $\Trunc{A}$, is a higher inductive type
    generated by the following constructors,
    \begin{itemize}
      \item an inclusion function $\trunc{\blank} : A \to \Trunc{A}$,
      \item for each $x, y : \Trunc{A}$, a path $\term{trunc}(x,y) : x \id_{\Trunc{A}} y$,
    \end{itemize}
    such that, given any type $B$ with
    \begin{itemize}
      \item a function $g : A \to B$,
      \item for each $x, y : B$, a path $\term{trunc*}(x,y) : x \id_{B} y$,
    \end{itemize}
    there is a unique function $f : \Trunc{A} \to B$ such that,
    \begin{itemize}
      \item $f(\trunc{a}) \equiv g(a)$
      \item for each $x, y : \Trunc{A}$, $\ap{f}{\term{trunc}(x,y)} \id_{B} \term{trunc*}(f(x),f(y))$.
    \end{itemize}
  \end{definition}

Another HIT that we use is the set-quotient $\quot{A}{R}$ which takes an $\hSet$ $A$ and a relation $R : A \to A \to
  \UU$. It has an inclusion of points $q : A \to \quot{A}{R}$, and adds paths between related pairs of elements
$\quotrel : R(x,y) \to q(x) \id_{\quot{A}{R}} q(y)$ (see~\cref{def:set-quot}).  We recall that the quotient is
\emph{effective} if $R$ is a prop-valued equivalence relation, that is, $R(x,y)$ holds iff $(q(x) \id_{\quot{A}{R}}
  q(y))$.

  \begin{definition}[Set Quotient]
    \label{def:set-quot}
    Given a type $A$ which is an $\hSet$, and a relation $R : A \to A \to \hProp$, the set-quotient $\quot{A}{R}$ is the
    higher inductive type generated by
    \begin{itemize}
      \item an inclusion function $q : A \to \quot{A}{R}$,
      \item for each $x, y : A$ such that $R(x,y)$, a path $q(x) \id_{\quot{A}{R}} q(y)$,
      \item a set truncation, for each $x, y : \quot{A}{R}$ and $r, s : x \id_{\quot{A}{R}} y$, we have $r \id s$,
    \end{itemize}
    with \review{an appropriate induction principle.}
  \end{definition}

\end{toappendix}

\subsection{Univalent Subuniverses}

Starting from a univalent universe which classifies all types, we want to define a subuniverse which classifies only
certain types, for example, types that satisfy some desired property. We use a prop-valued type family, that is, a
predicate on the universe, which picks out only those types, and collect them into a univalent subuniverse. Being
univalent ensures that the equality type of the ambient universe is reflected in the subuniverse.

\begin{definition}[Universe]
  A universe \`{a} la Tarski is given by the following pieces of data,
  \begin{itemize}
    \item a code $U : \UU$,
    \item a decoding type family $\El : U \to \UU$.
  \end{itemize}
  If $\El$ is univalent, $U$ is a univalent universe.
\end{definition}

\begin{definition}[Subuniverse]
  A universe predicate is a type family $P : \UU \to \UU$ whose fibers are prop-valued, that is, $P(X)$ is a proposition
  for every $X$. A subuniverse generated by $P$ is given by its total space $U \defeq \dsum{X:\UU}{P(X)}$ and $\El
  \defeq \fst$.
\end{definition}

\begin{proposition}[Univalent Subuniverse]
  Subuniverses generated by predicates are univalent.
\end{proposition}

\begin{appendixproof}
  Suppose $(U, \El) \defeq (\dsum{X:\UU}{P(X)}, \fst)$ is a subuniverse generated by a subtype $P : \UU \to \UU$. For
  any $X, Y : \UU$ such that $\phi : P(X)$ and $\psi : P(Y)$, we want to show that
  $\tptEqv{\fst} : (X,\phi) \id (Y,\psi) \to X \eqv Y$ is an equivalence. We construct
  $X \eqv Y \to (X,\phi) \id (Y,\psi)$ by $\ua$ and using the fact that $P(\blank)$ is a proposition. That it is an
  inverse follows by calculation using the appropriate computation rules.
\end{appendixproof}

The types we're interested in are the finite types. In constructive mathematics, the notion of finiteness is subtle. We
define a type to be (Bishop) finite if it is equal to some finite set, and we setup the predicate $\isFin$ to do so.
\todo{Cite Brent Yorgey's thesis?}

\begin{definition}[$\Fin$]
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by their cardinality. It is defined
  equivalently in two different ways,
  \begin{gather*}
    \begin{aligned}
      \Fin[n] & \defeq \dsum{k:\Nat}{k < n}
    \end{aligned}
    \qquad
    \begin{aligned}
       & \Fin[0] \defeq \bot                      \\
       & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a $\hSet$, and we use the definitions interchangeably.
\end{definition}

\begin{definition}[$\isFin$]
  We say that a type is finite if it is merely equal to $\Fin[n]$ for some $n$.
  \[
    \isFin[X] \defeq \dsum{n:\Nat}{\SubP{X}{\Fin[n]}}
  \]
  Note that the natural number $n$ need not be truncated, as justified below.
\end{definition}

\begin{proposition}
  For any type $X$, $\isFin[X]$ is a proposition.
\end{proposition}

\begin{appendixproof}
  Suppose we have $(n,\phi) : \isFin[X]$ and $(m,\psi) : \isFin[X]$, we need to show that $(n,\phi) \id (m,\psi)$. It is
  enough to show that $n \id m$. Since $\Nat$ is a set, this is a proposition, so we can use the induction principle of
  propositional truncation to eliminate to $n \id m$, applying it on $\phi$ and $\psi$ respectively. This gives us the
  equalities $X \id \Fin[n]$ and $X \id \Fin[m]$, which gives us $\Fin[n] \id \Fin[m]$, from which $n \id m$ follows by
  applying the first projection.
\end{appendixproof}

\begin{definition}
  The univalent subuniverse of \emph{all finite types} is given by
  \[
    \UFin \defeq \dsum{X:\UU}{\isFin[X]}.
  \]
  We write $F_{n} \defeq (\Fin[n], n, \trunc{\refl})$, for the image of the inclusion of $\Fin[n]$.
\end{definition}

We also consider a different way of thinking about $\UFin$, using the subuniverse $\BAut[T]$.

\begin{definition}[$\BAut$]
  The predicate $P(X) \defeq \Trunc[-1]{X \id T}$ picks out exactly those types that are merely equal to $T$, and this
  generates the subuniverse
  \[
    \BAut[T] \defeq \Sub{T}.
  \]
  We write $T_0 \defeq (T, \trunc{\refl_{T}})$ for the image of the inclusion of $T$ in $\BAut[T]$
\end{definition}

Using $\BAut$, we can talk about types that have a specified cardinality, for example, the subuniverse of 2-element sets
is given by $\BAut[\Bool]$. This has been used to construct the real projective spaces in
HoTT~\cite{buchholtzRealProjectiveSpaces2017}, and also to give the denotational semantics for a 1-bit reversible
language~\cite{caretteReversibleProgramsUnivalent2018}.

\begin{definition}
  For any $n : \Nat$, we define $\UFin[n] \defeq \BAut[\Fin[n]]$ to be the univalent subuniverse of $n$-element sets.
  Note that, $\UFin$ can be equivalently seen as the collection of all types of finite cardinality, that is, $\UFin \eqv
    \displaystyle\sqcup_{n:\Nat} \UFin[n]$.
\end{definition}

Since $\BAut[T]$ is a univalent subuniverse, we can characterise its path space. The intuition is that $\BAut[T]$ only
has one point $T_0$, and 1-paths $T_0 \id_{\BAut[T]} T_0$, that is, loops, and higher paths between these loops. The
type of loops on $T_0$, $\loopspace[\BAut[T],T_{0}]$ is shown to be equivalent to $\Aut[T] \defeq T \eqv T$, which is
the automorphism group of $T$.

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
    \item For any $T : \UU$, $\BAut[T]$ is 0-connected.
    \item For any $T : \UU$, \( \loopspace[\BAut[T],T_{0}] \eqv \Aut[T] \). \label{lem:loop-deloop}
  \end{enumerate}
\end{proposition}

\begin{appendixproof}
  We need to show that the equality type of $\BAut[T]$ is an $n$-type. Assume $X, Y : \BAut[T]$. Since $\BAut[T]$ is a
  univalent subuniverse, we have $(X \id Y) \eqv (\fst(X) \eqv \fst(Y))$. Note that being an $n$-type is a proposition.
  Since $T$ is an $n$-type, and $\fst(X)$ and $\fst(Y)$ are merely equal to $T$, they're also $n$-types. It follows that
  $\fst(X) \eqv \fst(Y)$ is an $n$-type, and hence $X \id Y$ is an $n$-type.

  Since $\BAut[T]$ is a univalent universe, it follows that
  \[
    (T_{0} \id_{\BAut[T]} T_{0}) \eqv (\fst(T_{0}) \eqv \fst(T_{0})) \equiv (T \eqv T) \equiv \Aut[T].
  \]
\end{appendixproof}

\begin{corollary}
  $\UFin$ is a pointed, connected, 1-groupoid. For every $n:\Nat$, \( \loopspace[\UFin[n],F_{n}] \eqv \Aut[\Fin[n]] \).
\end{corollary}

The loopspace of a pointed type automatically has the structure of a group, that is, with $\refl_{\pt}$ for the neutral
element, path composition for the group multiplication, and path inverse for the group inverse operation. The group
axioms are then given by the higher paths corresponding to groupoid laws. We have shown that loops in $\UFin$ exactly
encode the automorphism group $\Aut[\Fin[n]]$ for every $n$. This is a general technique called delooping, where a group
can be identified with a 1-object groupoid, internally in HoTT.~\footnote{This technique also allows defining higher
groups in HoTT, as done in~\cite{buchholtzHigherGroupsHomotopy2018}.}

% \[
%   \begin{tikzcd}
%     F_{0}
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
%     & F_{1}
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
%     & F_{2}
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=8em, in=125, out=55]
%     & \ldots
%     & F_{n}
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=4em, in=115, out=65]
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=8em, in=125, out=55]
%     \arrow[""{anchor=center, inner sep=0}, no head, loop, distance=12em, in=135, out=45]
%     & \ldots
%   \end{tikzcd}
% \]

\subsection{Rig structure}~\label{subsec:rig}

The groupoid $\UFin$ has two symmetric monoidal structures, the additive and the multiplicative one, and the
multiplicative tensor product distributes over the additive one. To construct these, we first define some equivalences
on $\Fin$, and some general type isomorphisms.

\begin{proposition}
  For any $n, m : \Nat$,
  \begin{gather*}
    \begin{aligned}
      \Fin[0]                & \eqv \bot        \\
      \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m] \\
    \end{aligned}
    \qquad
    \begin{aligned}
      \Fin[1]                & \eqv \top        \\
      \Fin[n] \times \Fin[m] & \eqv \Fin[n * m] \\
    \end{aligned}
  \end{gather*}
  and for any types $X, Y, Z$,
  \begin{gather*}
    \begin{aligned}
      \bot \sqcup X         & \eqv X                     \\
      X \sqcup \bot         & \eqv X                     \\
      (X \sqcup Y) \sqcup Z & \eqv X \sqcup (Y \sqcup Z) \\
      X \sqcup Y            & \eqv Y \sqcup Y            \\
    \end{aligned}
    \qquad
    \begin{aligned}
      \top \times X         & \eqv X                                \\
      X \times \top         & \eqv X                                \\
      (X \times Y) \times Z & \eqv X \times (Y \times Z)            \\
      X \times Y            & \eqv Y \times X                       \\
      X \times \bot         & \eqv \bot                             \\
      X \times (Y \sqcup Z) & \eqv (X \times Y) \sqcup (X \times Z)
    \end{aligned}
  \end{gather*}
\end{proposition}

We lift these equivalences to $\UFin$ giving it the additive and multiplicative symmetric monoidal structures, $(F_0,
  \sqcup)$ and $(F_1, \times)$, with corresponding natural isomorphisms $\lambda_{X}$, $\rho_{X}$, $\alpha_{X,Y,Z}$, and
the braiding isomorphism $\mathcal{B}_{X,Y}$. These isomorphisms satisfy the Mac Lane coherence conditions for symmetric
monoidal categories, that is, the triangle, pentagon, and hexagon identities, and the symmetry of the braiding, upto
2-paths in $\UFin$.

\begin{toappendix}
  \begin{definition}
    \begin{align*}
      O                 & \defeq F_{0}                                       \\
      X \oplus Y        & \defeq X \sqcup Y                                  \\
      \lambda_{X}       & : O \oplus X \eqv X                                \\
      \rho_{X}          & : X \oplus O \eqv X                                \\
      \alpha_{X,Y,Z}    & : (X \oplus Y) \oplus Z \eqv X \oplus (Y \oplus Z) \\
      \mathcal{B}_{X,Y} & : X \oplus Y \eqv Y \oplus X
    \end{align*}
  \end{definition}
  \begin{proposition}
    % https://q.uiver.app/?q=WzAsNCxbMCwwLCIoWCBcXG9wbHVzIEkpIFxcb3BsdXMgWSJdLFsyLDAsIlggXFxvcGx1cyAoSSBcXG9wbHVzIFkpIl0sWzEsMSwiWCBcXG9wbHVzIFkiXSxbMCwxXSxbMCwxLCJcXGFscGhhX3tYLEksWX0iXSxbMCwyLCJcXHJob197WH0gXFxvcGx1cyAxX3tZfSIsMl0sWzEsMiwiMV97WH0gXFxvcGx1cyBcXGxhbWJkYV97WX0iXSxbNSw2LCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}
        {(X \oplus I) \oplus Y} && {X \oplus (I \oplus Y)} \\
        {} & {X \oplus Y}
        \arrow["{\alpha_{X,I,Y}}", from=1-1, to=1-3]
        \arrow[""{name=0, anchor=center, inner sep=0}, "{\rho_{X} \oplus 1_{Y}}"', from=1-1, to=2-2]
        \arrow[""{name=1, anchor=center, inner sep=0}, "{1_{X} \oplus \lambda_{Y}}", from=1-3, to=2-2]
        \arrow["\id", Rightarrow, draw=none, from=0, to=1]
      \end{tikzcd}\]
    % https://q.uiver.app/?q=WzAsNSxbMCwxLCIoKFcgXFxvcGx1cyBYKSBcXG9wbHVzIFkpIFxcb3BsdXMgWiJdLFsxLDAsIihXIFxcb3BsdXMgWCkgXFxvcGx1cyAoWSBcXG9wbHVzIFopIl0sWzIsMSwiVyBcXG9wbHVzIChYIFxcb3BsdXMgKFkgXFxvcGx1cyBaKSkiXSxbMiwzLCJXIFxcb3BsdXMgKChYIFxcb3BsdXMgWSkgXFxvcGx1cyBaKSJdLFswLDMsIihXIFxcb3BsdXMgKFggXFxvcGx1cyBZKSkgXFxvcGx1cyBaIl0sWzAsMSwiXFxhbHBoYV97VyBcXG9wbHVzIFgsIFksIFp9Il0sWzEsMiwiXFxhbHBoYV97VyxYLFkgXFxvcGx1cyBafSJdLFszLDIsIjFfe1d9IFxcb3BsdXMgXFxhbHBoYV97WCxZLFp9IiwyXSxbMCw0LCJcXGFscGhhX3tXLFgsWX0gXFxvcGx1cyAxX3tafSIsMl0sWzQsMywiXFxhbHBoYV97VyxYIFxcb3BsdXMgWSxafSIsMl0sWzAsMiwiXFxpZCIsMSx7Im9mZnNldCI6NSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}
        & {(W \oplus X) \oplus (Y \oplus Z)} \\
        {((W \oplus X) \oplus Y) \oplus Z} && {W \oplus (X \oplus (Y \oplus Z))} \\
        \\
        {(W \oplus (X \oplus Y)) \oplus Z} && {W \oplus ((X \oplus Y) \oplus Z)}
        \arrow["{\alpha_{W \oplus X, Y, Z}}", from=2-1, to=1-2]
        \arrow["{\alpha_{W,X,Y \oplus Z}}", from=1-2, to=2-3]
        \arrow["{1_{W} \oplus \alpha_{X,Y,Z}}"', from=4-3, to=2-3]
        \arrow["{\alpha_{W,X,Y} \oplus 1_{Z}}"', from=2-1, to=4-1]
        \arrow["{\alpha_{W,X \oplus Y,Z}}"', from=4-1, to=4-3]
        \arrow["\id"{description}, shift right=5, draw=none, from=2-1, to=2-3]
      \end{tikzcd}\]
    % https://q.uiver.app/?q=WzAsNixbMSwwLCJYIFxcb3BsdXMgKFkgXFxvcGx1cyBaKSJdLFswLDEsIihYIFxcb3BsdXMgWSkgXFxvcGx1cyBaIl0sWzAsMiwiKFkgXFxvcGx1cyBYKSBcXG9wbHVzIFoiXSxbMSwzLCJZIFxcb3BsdXMgKFggXFxvcGx1cyBaKSJdLFsyLDIsIlkgXFxvcGx1cyAoWiBcXG9wbHVzIFgpIl0sWzIsMSwiKFkgXFxvcGx1cyBaKSBcXG9wbHVzIFgiXSxbMSwwLCJcXGFscGhhX3tYLFksWn0iXSxbMSwyLCJcXG1hdGhjYWx7Qn1fe1gsWX0gXFxvcGx1cyAxX3tafSIsMl0sWzIsMywiXFxhbHBoYV97WSxYLFp9IiwyXSxbMyw0LCIxX3tZfSBcXG9wbHVzIFxcbWF0aGNhbHtCfV97WCxafSIsMl0sWzUsNCwiXFxhbHBoYV97WSxaLFh9Il0sWzAsNSwiXFxtYXRoY2Fse0J9X3tYLFkgXFxvcGx1cyBafSJdLFs3LDEwLCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}
        & {X \oplus (Y \oplus Z)} \\
        {(X \oplus Y) \oplus Z} && {(Y \oplus Z) \oplus X} \\
        {(Y \oplus X) \oplus Z} && {Y \oplus (Z \oplus X)} \\
        & {Y \oplus (X \oplus Z)}
        \arrow["{\alpha_{X,Y,Z}}", from=2-1, to=1-2]
        \arrow[""{name=0, anchor=center, inner sep=0}, "{\mathcal{B}_{X,Y} \oplus 1_{Z}}"', from=2-1, to=3-1]
        \arrow["{\alpha_{Y,X,Z}}"', from=3-1, to=4-2]
        \arrow["{1_{Y} \oplus \mathcal{B}_{X,Z}}"', from=4-2, to=3-3]
        \arrow[""{name=1, anchor=center, inner sep=0}, "{\alpha_{Y,Z,X}}", from=2-3, to=3-3]
        \arrow["{\mathcal{B}_{X,Y \oplus Z}}", from=1-2, to=2-3]
        \arrow["\id", Rightarrow, draw=none, from=0, to=1]
      \end{tikzcd}\]
    % https://q.uiver.app/?q=WzAsMyxbMCwwLCJYIFxcb3BsdXMgWSJdLFsyLDAsIlggXFxvcGx1cyBZIl0sWzEsMSwiWSBcXG9wbHVzIFgiXSxbMCwxLCIxX3tYIFxcb3BsdXMgWX0iLDAseyJsZXZlbCI6Miwic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFswLDIsIlxcbWF0aGNhbHtCfV97WCxZfSIsMl0sWzIsMSwiXFxtYXRoY2Fse0J9X3tZLFh9IiwyXSxbNCw1LCJcXGlkIiwwLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}
        {X \oplus Y} && {X \oplus Y} \\
        & {Y \oplus X}
        \arrow["{1_{X \oplus Y}}", Rightarrow, no head, from=1-1, to=1-3]
        \arrow[""{name=0, anchor=center, inner sep=0}, "{\mathcal{B}_{X,Y}}"', from=1-1, to=2-2]
        \arrow[""{name=1, anchor=center, inner sep=0}, "{\mathcal{B}_{Y,X}}"', from=2-2, to=1-3]
        \arrow["\id", Rightarrow, draw=none, from=0, to=1]
      \end{tikzcd}\]
  \end{proposition}
\end{toappendix}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End: