### 2021-10-29 Pi/paper/applications.tex
Reorganise applications section

\section{Applications to Reversible Circuits}
\label{sec:applications}

Using our semantics, we can normalise, synthesise, prove equivalence, and reason about~$\PiLang$ programs.

\[
  \begin{tikzpicture}[->,>=stealth',auto,thick,node distance=1em]
    % nodes
    \node (A) at (0, 2) {\Afun{reversibleOr1}};
    \node (B) at (0, 0) {\Afun{reversibleOr2}};
    \node (C) at (4, 1) {\Afun{reversibleOrNorm}};
    \node (D) at (9, 1) {\AgdaSymbol{(}\AgdaNumber{0 5 6 7 4 1 2 3}\AgdaSymbol{)}};
    % arrows
    \path
    (A) edge[bend left] node [above left] { $\strut\normt_{1}$ } (C)
    (B) edge[bend right] node [below left] { $\strut\normt_{1}$ } (C)
    (D) edge node[above] { $\strut\term{quote}$ } (C);
  \end{tikzpicture}
\]

% \[
%   \begin{tabular}{>{$}c<{$} @{\hspace{5em}} >{$}c<{$}}
%     \tikzmarknode{A}{\Afun{reversibleOr1}} & \\
%                                            & \tikzmarknode{C}{\Afun{reversibleOrNorm}} \\
%     \tikzmarknode{B}{\Afun{reversibleOr2}} &
%   \end{tabular}
%   % \tikzmarknode{D}{4}
%   \begin{tikzpicture}[overlay, remember picture, shorten <=1mm,
%     nodes={inner sep=1pt, align=center, font=\footnotesize}]
%     \draw (A.south) -- ++ (-1,-1) node[below] {2 bits};
%     \draw (B.south) -- ++ (-.5,-1) node[below] {2 bits};
%     % \draw (D.south) -- ++ (.5,-1) node[below] {3 bits};
%   \end{tikzpicture}
% \]

\subsection{Normalisation}

Normalisation involves translating $\PiLang$ programs to $\PiHatLang$, computing a permutation, and quoting back to
$\PiPlusLang$. This happends in several steps -- the evaluation step goes from $\PiLang$ to $\PiPlusLang$ reducing
$\times$ to $+$ and translating the corresponding combinators, then from $\PiPlusLang$ to $\PiHatLang$ reducing to
adjacent transpositions, then finally to $\UFin$ through words in $\Sn$ and Lehmer codes, and then, the quoting step
goes all the way back until $\PiPlusLang$. We define this $\normt$ function below using our previous evaluation
$\gdenot{\blank}$.

\begin{definition}[Normalisation of $\PiLang$ programs]
  \begin{gather*}
    \begin{aligned}
      \normt_{0} & : U \to \UPlus                                            \\
      \normt_{0} & = \quotep_{0} \comp \quoteh_{0} \comp \gdenot{\blank}_{0}
    \end{aligned}
    \qquad
    \begin{aligned}
      \normt_{1} & : (c : X \iso Y) \to \normt_{0}(X) \iso \normt_{0}(Y)     \\
      \normt_{1} & = \quotep_{1} \comp \quoteh_{1} \comp \gdenot{\blank}_{1}
    \end{aligned}
  \end{gather*}
\end{definition}

\subsection{Equivalence}

By normalisation, we also get a decision procedure for program equivalence -- two circuits are equivalent iff they
represent the same permutation, or equivalently, they have the same normal form. Recall the specification of reversible
disjunction from \cref{sec:examples}, the two Qiskit circuits for implementing it, and the corresponding $\PiLang$
definitions \Afun{reversibleOr1} and \Afun{reversibleOr2}. The type $\mathbb{B}\ 3$ normalises to
$\mathbb{8}+ \equiv \mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + \mathbb{0})))))))$,
and normal forms for both circuits compute to the following, establishing their equivalence:

\medskip
\resetnormtwo{}

\subsection{Synthesis}

Instead of manually producing $\Pi$-programs to implement the reversible disjunction specification, it is possible to
simply quote the desired permutation, synthesising a $\PiLang$ program. We use a function $\term{lookup}$ to write
tabulated permutations for readability.

\medskip
\resetperm{}

\noindent
The permutation uses the canonical encoding of sequences of bits as natural numbers (e.g., (\textsf{false},
\textsf{true},\textsf{true}) is encoded as 011 or 3). The second entry maps index 1 (= 001) to the value 5 (= 101) which
states that since one of the right bits is set in 001 then the leftmost bit in the output is set. Quoting this
permutation generates the same normalised program matching the desired structured types.

\subsection{Reasoning}

More general, user-guided, reasoning can be done using the sound and complete level-2 combinators to rewrite $\PiLang$
programs.

% \paragraph*{Program synthesis.} The NbE process embodies a quoting mechanism that synthesizes programs from
% permutations. Indeed, instead of writing a program for \Afun{reset 2}, one could simply specify the desired permutation
% as:

%%\begin{minipage}{.65\textwidth}
%%  \PiRESET{}
%%\end{minipage}
%%\begin{minipage}{.30\textwidth}
%%  \begin{center}
%%  \resizebox{0.5\textwidth}{!}{\input{diagrams/reset.tikz}}
%%  \end{center}
%%\end{minipage}

%% \resettwo{}

% \noindent The syntax will be explained in detail in the next section and the full definitions of the helpers are
% provided in the supplementary material. For now, it is sufficient to know that there is some program that implements the
% reversible function of interest and that applying \Afun{reset} to 2 produces \verb|reversibleOr2| from the introduction.

% In principle, the normalised program can be produced following two strategies: (i) by repeatedly applying the rewrite
% rules of our calculus of reversible functions (explained in Sec.~\ref{sec:reversibletwo}), or (ii) in the case above by
% using a NbE process that evaluates the program to a permutation on a finite set of 8
% elements and reifies that permutation back to a program. The key idea of the NbE process is a systematic way to express
% permutations as sequences of adjacent swaps as illustrated in the following small example where the permutation on the
% left is compiled to the sequence of four adjacent transpositions on the right:

% \note{Motivation: There are two reversible circuits which describe the following permutation. They can be shown to be
%   equal using the 2-combinators.}

% \[
%   \begin{tikzpicture}
%     \begin{knot}[clip width=5]
%       \filldraw (0,5) circle (2pt) node[above] {0};
%       \filldraw (1,5) circle (2pt) node[above] {1};
%       \filldraw (2,5) circle (2pt) node[above] {2};
%       \filldraw (3,5) circle (2pt) node[above] {3};
%       \filldraw (4,5) circle (2pt) node[above] {4};
%       \filldraw (0,0) circle (2pt) node[below] {1};
%       \filldraw (1,0) circle (2pt) node[below] {4};
%       \filldraw (2,0) circle (2pt) node[below] {0};
%       \filldraw (3,0) circle (2pt) node[below] {3};
%       \filldraw (4,0) circle (2pt) node[below] {2};
%       \strand (0,5) .. controls (0.5,0.5) and (1.5,3.5) .. (2,0);
%       \strand (1,5) .. controls (0.75,0.5) and (0.25,3.5) .. (0,0);
%       \strand (2,5) .. controls (2.5,2.5) and (3.5,1.5) .. (4,0);
%       \strand (3,5) .. controls (4.5,2.5) and (4,1.5) .. (3,0);
%       \strand (4,5) .. controls (3.5,2.5) and (1.5,2.5) .. (1,0);
%       \flipcrossings{4,5};
%     \end{knot}
%   \end{tikzpicture}
% \]

% \paragraph*{Program equivalence.} The permutation above reveals another way to think about the desired program: it is a
% special addition circuit that keeps 0 and 4 fixed but otherwise adds 4 modulo 8 to its input. From this specification,
% one can use a standard synthesis algorithm for reversible circuits~\cite{10.1145/775832.775915} to generate the
% following program:

% \adder{}

% \noindent The \Afun{adder3} program looks nothing like the original \Afun{reset 2} program and yet they both have the
% same normal form thus establishing their equivalence. The reader can check that this circuit is the same as
% \verb|reversibleOr1| from the introduction.

% Now imagine we want to write the following reversible function:

% 0 -> 0
% 8 -> 8
% n -> n + 8 `mod` 16

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End: