### 2021-10-30 Pi/paper/applications.tex
Progress

\section{Applications to Reversible Circuits}
\label{sec:applications}

Using our semantics, we can normalise, synthesise, prove equivalence, and reason about~$\PiLang$ programs.

% \[
%   \begin{tikzpicture}[->,>=stealth',auto,thick,node distance=1em]
%     % nodes
%     \node (A) at (0, 2) {\Afun{reversibleOr1}};
%     \node (B) at (0, 0) {\Afun{reversibleOr2}};
%     \node (C) at (4, 1) {\Afun{reversibleOrNorm}};
%     \node (D) at (9, 1) {\AgdaSymbol{(}\AgdaNumber{0 5 6 7 4 1 2 3}\AgdaSymbol{)}};
%     % arrows
%     \path
%     (A) edge[bend left] node [above left] { $\strut\normt_{1}$ } (C)
%     (B) edge[bend right] node [below left] { $\strut\normt_{1}$ } (C)
%     (D) edge node[above] { $\strut\term{quote}$ } (C);
%   \end{tikzpicture}
% \]

% \[
%   \begin{tabular}{>{$}c<{$} @{\hspace{5em}} >{$}c<{$}}
%     \tikzmarknode{A}{\Afun{reversibleOr1}} & \\
%                                            & \tikzmarknode{C}{\Afun{reversibleOrNorm}} \\
%     \tikzmarknode{B}{\Afun{reversibleOr2}} &
%   \end{tabular}
%   % \tikzmarknode{D}{4}
%   \begin{tikzpicture}[overlay, remember picture, shorten <=1mm,
%     nodes={inner sep=1pt, align=center, font=\footnotesize}]
%     \draw (A.south) -- ++ (-1,-1) node[below] {2 bits};
%     \draw (B.south) -- ++ (-.5,-1) node[below] {2 bits};
%     % \draw (D.south) -- ++ (.5,-1) node[below] {3 bits};
%   \end{tikzpicture}
% \]

\subsection{Normalisation}

Normalisation happens by evaluating a combinator to a permutation and then reifying it back. The evaluation step goes
from $\PiLang$ to $\PiPlusLang$ reducing $\times$ to $+$ and translating the corresponding combinators, then from
$\PiPlusLang$ to $\PiHatLang$ reducing them to natural numbers and adjacent transpositions, then finally to $\UFin$
through words in $\Sn$ and Lehmer codes. Then, the quotation step goes all the way back until $\PiPlusLang$. We define
this composite $\normt$ function below using our previous evaluation $\gdenot{\blank}$.

\begin{definition}[Normalisation of $\PiLang$ programs]
  \begin{gather*}
    \begin{aligned}
      \normt_{0} & : U \to \UPlus                                            \\
      \normt_{0} & = \quotep_{0} \comp \quoteh_{0} \comp \gdenot{\blank}_{0}
    \end{aligned}
    \qquad
    \begin{aligned}
      \normt_{1} & : (c : X \iso Y) \to \normt_{0}(X) \iso \normt_{0}(Y)     \\
      \normt_{1} & = \quotep_{1} \comp \quoteh_{1} \comp \gdenot{\blank}_{1}
    \end{aligned}
  \end{gather*}
\end{definition}

As an example, consider the \Afun{reversibleOr1} circuit introduced in~\cref{sec:pi}. After applying $\normt_{1}$, the
type $\mathbb{B}\ 3$ normalises to $\mathbb{8}+ \equiv \mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} +
  (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + (\mathbb{1} + \mathbb{0})))))))$, and the circuit reduces to the following
normal form:

\medskip
\resetnormtwo{}

\subsection{Equivalence}

By normalisation, we also get a decision procedure for program equivalence -- two circuits are equivalent iff they
represent the same permutation, or equivalently, they have the same normal form. Recall that in~\cref{sec:examples} we
introduced two different ways of computing reversible disjunction using two Qiskit circuit implementations, and then
in~\cref{sec:pi} showed the corresponding $\PiLang$ definitions: \Afun{reversibleOr1} and \Afun{reversibleOr2}.

After applying $\normt_{1}$ to each circuit, we verify that they both reduce to the same normal form
\Afun{reversibleOrNorm}, along with a 2-combinator witnessing their equivalence with the normal form -- by composing
these, we get the 2-combinator witnessing their equivalence.

\subsection{Synthesis}

Instead of manually producing $\PiLang$-programs to implement the reversible disjunction specification, it is possible
to simply write down the permutation directly and then quote it, synthesising a $\PiLang$ program. We use a function
$\term{lookup}$ to write tabulated permutations for readability.

\medskip
\resetperm{}

\noindent
The permutation uses the canonical encoding of sequences of bits as natural numbers, e.g., {(\textsf{false},
\textsf{true},\textsf{true})} is encoded as 011 or 3. The second entry in the $\term{lookup}$ table maps index 1 (= 001)
to the value 5 (= 101), following the reversible disjunction specification (recall that
$\mathsf{reversibleOr}(h,b_1,b_2) = (h \,\underline{\vee}\, (b_1 \vee b_2), ~b_1, ~b_2)$ where~$\vee$ is disjunction
and~$\underline{\vee}$ is exclusive-disjunction). Quoting this permutation generates the same normalised program
\Afun{reversibleOrNorm}, matching the desired structured type of a 1-combinator on $\mathbb{8}+$.

\subsection{Reasoning}

In our development, we described a number of different representations of permutations: $\PiLang$ programs, lists of
transpositions, Lehmer codes, and automorphisms of finite sets. The equivalences between these representations allows us
to transfer theorems about permutations, proving them on the representations most suited for the particular task and
getting others for free.

As an example, consider the following problem. Toffoli gate on 3 bits can be extended to 4 bits, inserting a wire in
four possible positions. The function of such gate is then identity on the newly inserted bit, and controlled-controlled
not on the three remaining ones.
\medskip
\extendedToffoli{}
\baselineskip
On the other hand, it is possible to construct a proper 4-bit toffoli gate:
\medskip
\toffoli{}
\baselineskip
Now, the question we want to answer is: is it possible to compose some number of (extended) 3-bit toffoli gates in some way, so as to replicate the
behaviour of a 4-bit toffoli gate?

Solving this problem using the circuit representation is difficult. However, we can observe that 3-bit toffoli gate
implements an even permutation, that is, a permutation consisting of an even number of transpositions. On the other
hand, 4-bit toffoli gate is an odd permutation. It is an easily proven lemma that composition of even permutations
always produces an even permutation, which answers the question to the negative.

We define a function $\parity : (c : X \iso Y) \to \Bool$ to compute a parity of a permutation defined by a combinator
$c$. It is implemented as the transformation from a $\PiLang$ program into a list of transposition, using $\evalh_{1}$
defined in~\cref{sec:equivalence}, and computing the parity on this representation, which means simply checking the
parity of the length of the list.

\begin{propositionrep}
  \label{prop:parity-preserved}
  For two $\PiLang$ circuits $c_1, c_2$, if there exists a 2-combinator between them, then $\parity c_1 = \term{parity} c_2$. 
\end{propositionrep}
\begin{proof}
  First, it is easy to check that Coxeter relations preserve parity of the list. Second, if $c_1, c_2$ are related by a
  2-combinator, then their list-of-transpositions representations are related in the Coxeter relation. Composing these
  two facts completes the proof.
\end{proof}

Using ~\cref{prop:parity-preserved} and showing that any composition of even cicuits yields an even ciruit, we can
answer the question by computing the parity of the \Afun{toffoli₄} and all of the \Afun{toffoli₃} gates.

Thus, we formulated the theorem using the representation on lists of transpositions, and transferred it to circuits using
the equivalences proven in the previous sections.

% More general, user-guided, reasoning can be done using the sound and complete level-2 combinators to rewrite $\PiLang$
% programs.

% \paragraph*{Program synthesis.} The NbE process embodies a quoting mechanism that synthesizes programs from
% permutations. Indeed, instead of writing a program for \Afun{reset 2}, one could simply specify the desired permutation
% as:

%%\begin{minipage}{.65\textwidth}
%%  \PiRESET{}
%%\end{minipage}
%%\begin{minipage}{.30\textwidth}
%%  \begin{center}
%%  \resizebox{0.5\textwidth}{!}{\input{diagrams/reset.tikz}}
%%  \end{center}
%%\end{minipage}

%% \resettwo{}

% \noindent The syntax will be explained in detail in the next section and the full definitions of the helpers are
% provided in the supplementary material. For now, it is sufficient to know that there is some program that implements the
% reversible function of interest and that applying \Afun{reset} to 2 produces \verb|reversibleOr2| from the introduction.

% In principle, the normalised program can be produced following two strategies: (i) by repeatedly applying the rewrite
% rules of our calculus of reversible functions (explained in Sec.~\ref{sec:reversibletwo}), or (ii) in the case above by
% using a NbE process that evaluates the program to a permutation on a finite set of 8
% elements and reifies that permutation back to a program. The key idea of the NbE process is a systematic way to express
% permutations as sequences of adjacent swaps as illustrated in the following small example where the permutation on the
% left is compiled to the sequence of four adjacent transpositions on the right:

% \note{Motivation: There are two reversible circuits which describe the following permutation. They can be shown to be
%   equal using the 2-combinators.}

% \[
%   \begin{tikzpicture}
%     \begin{knot}[clip width=5]
%       \filldraw (0,5) circle (2pt) node[above] {0};
%       \filldraw (1,5) circle (2pt) node[above] {1};
%       \filldraw (2,5) circle (2pt) node[above] {2};
%       \filldraw (3,5) circle (2pt) node[above] {3};
%       \filldraw (4,5) circle (2pt) node[above] {4};
%       \filldraw (0,0) circle (2pt) node[below] {1};
%       \filldraw (1,0) circle (2pt) node[below] {4};
%       \filldraw (2,0) circle (2pt) node[below] {0};
%       \filldraw (3,0) circle (2pt) node[below] {3};
%       \filldraw (4,0) circle (2pt) node[below] {2};
%       \strand (0,5) .. controls (0.5,0.5) and (1.5,3.5) .. (2,0);
%       \strand (1,5) .. controls (0.75,0.5) and (0.25,3.5) .. (0,0);
%       \strand (2,5) .. controls (2.5,2.5) and (3.5,1.5) .. (4,0);
%       \strand (3,5) .. controls (4.5,2.5) and (4,1.5) .. (3,0);
%       \strand (4,5) .. controls (3.5,2.5) and (1.5,2.5) .. (1,0);
%       \flipcrossings{4,5};
%     \end{knot}
%   \end{tikzpicture}
% \]

% \paragraph*{Program equivalence.} The permutation above reveals another way to think about the desired program: it is a
% special addition circuit that keeps 0 and 4 fixed but otherwise adds 4 modulo 8 to its input. From this specification,
% one can use a standard synthesis algorithm for reversible circuits~\cite{10.1145/775832.775915} to generate the
% following program:

% \adder{}

% \noindent The \Afun{adder3} program looks nothing like the original \Afun{reset 2} program and yet they both have the
% same normal form thus establishing their equivalence. The reader can check that this circuit is the same as
% \verb|reversibleOr1| from the introduction.

% Now imagine we want to write the following reversible function:

% 0 -> 0
% 8 -> 8
% n -> n + 8 `mod` 16

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End: